### 2024-12-16 landscapemodels.Rmd
analysis

---
title: "Landscape models"
output: html_notebook
---

```{r setup, echo = FALSE}
library(tidyverse)
library(performance)
library(psych)

wfs <- readRDS("data/maxwfsresults_20241209.rds")
lsm <- read.csv("data/lsm_reduced.csv") %>% 
  mutate(hu = gsub("hu", "", hu))

wfs_lsm <- wfs %>% left_join(lsm, by = "hu") %>%
  select(c(hu, huc2, name, margin.ha, lu_ta,
           pollinators_base:sed_export_base, 
           delta_pollinators:delta_sed, 
           fields_max, field.ha, margin.ha, 
           lu_shdi:ntr_frac_mn)) %>%
  mutate(prn_ann_ratio = log(prn_agr_pland/ann_agr_pland),
         prn_pagr = 100*(prn_agr_pland/(ann_agr_pland+prn_agr_pland)),
         field_pland = field.ha/lu_ta,
         margin_dens = margin.ha/lu_ta)

scaled.data <- wfs_lsm %>% mutate(across(lu_shdi:field_pland, ~scale(.x)[,1]))

```

## Correlation between independent variables.

I selected a subset of variables that aren't strongly correlated and that represent potential drivers of ecosystem service patterns.

```{r correlation, echo = FALSE}

pairs.panels(wfs_lsm %>% 
               select(ntr_pland, prn_agr_pland, aglu_iji,
                      lu_shdi, aglu_frac_mn, margin_dens), 
             cex.cor = 1, cex.labels=1, cex=0.5)

```

## Pollinator models

### Pollinator delta

```{r pollinator_delta}

pol.dmod0 <- lm(delta_pollinators ~ ntr_pland + margin_dens + prn_agr_pland, data = scaled.data)
pol.dmod1 <- lm(delta_pollinators ~ ntr_pland + aglu_iji + prn_agr_pland + aglu_frac_mn + margin_dens + lu_shdi, data = scaled.data)
pol.dmod2 <- lm(delta_pollinators ~ ntr_pland + I(ntr_pland^2) + aglu_iji + prn_agr_pland + aglu_frac_mn + margin_dens + lu_shdi, data = scaled.data)

performance::compare_performance(pol.dmod0, pol.dmod1, pol.dmod2)

summary(pol.dmod2)

```

## Seasonal water yield

### Seasonal water yield delta

```{r swy_delta}

swy.dmod0 <- lm(delta_qb ~ ntr_pland + margin_dens + prn_agr_pland, data = scaled.data)
swy.dmod1 <- lm(delta_qb ~ ntr_pland + aglu_iji + prn_agr_pland + aglu_frac_mn + margin_dens + lu_shdi, data = scaled.data)
swy.dmod2 <- lm(delta_qb ~ ntr_pland + I(ntr_pland^2) + aglu_iji + prn_agr_pland + aglu_frac_mn + margin_dens + lu_shdi, data = scaled.data)

performance::compare_performance(swy.dmod0, swy.dmod1, swy.dmod2)

summary(swy.dmod2)

```

## Nutrient delivery ratio

### Nutrient delivery ratio delta

```{r ndr_delta}

ndr.dmod0 <- lm(delta_n_export ~ ntr_pland + margin_dens + prn_agr_pland, data = scaled.data)
ndr.dmod1 <- lm(delta_n_export ~ ntr_pland + aglu_iji + prn_agr_pland + aglu_frac_mn + margin_dens + lu_shdi, data = scaled.data)
ndr.dmod2 <- lm(delta_n_export ~ ntr_pland + I(ntr_pland^2) + aglu_iji + prn_agr_pland + aglu_frac_mn + margin_dens + lu_shdi, data = scaled.data)

performance::compare_performance(ndr.dmod0, ndr.dmod1, ndr.dmod2)

summary(ndr.dmod2)

```

## Sediment delivery ratio

### Sediment delivery ratio delta

```{r sdr_delta}

sdr.dmod0 <- lm(delta_sed ~ ntr_pland + margin_dens + prn_agr_pland, data = scaled.data)
sdr.dmod1 <- lm(delta_sed ~ ntr_pland + aglu_iji + prn_agr_pland + aglu_frac_mn + margin_dens + lu_shdi, data = scaled.data)
sdr.dmod2 <- lm(delta_sed ~ ntr_pland + I(ntr_pland^2) + aglu_iji + prn_agr_pland + aglu_frac_mn + margin_dens + lu_shdi, data = scaled.data)

performance::compare_performance(sdr.dmod0, sdr.dmod1, sdr.dmod2)

summary(sdr.dmod2)

```