### 2024-12-27 landscapemodels_gam.Rmd
more analysis attempts

---
title: "Landscape models with Generalized Additive Models"
---

```{r setup, echo = FALSE}
library(tidyverse)
library(performance)
library(psych)
library(mgcv)
library(ggeffects)
library(piecewiseSEM)
source("R/utils-plot.r")

wfs <- readRDS("data/maxwfsresults_20241217_fix.rds")
lsm <- read.csv("data/lsm_reduced.csv") %>% 
  mutate(hu = gsub("hu", "", hu))

wfs_lsm <- wfs %>% left_join(lsm, by = "hu") %>%
  select(c(hu, huc2, name, margin.ha, lu_ta,
           pollinators_base:sed_export_base, 
           pollinators_max:sed_export_max, 
           delta_pollinators:delta_sed, 
           fields_max, field.ha, margin.ha, 
           lu_shdi:ntr_frac_mn)) %>%
  mutate(prn_ann_ratio = log(prn_agr_pland/ann_agr_pland),
         prn_pagr = 100*(prn_agr_pland/(ann_agr_pland+prn_agr_pland)),
         field_pland = field.ha/lu_ta,
         margin_dens = margin.ha/lu_ta)

scaled.data <- wfs_lsm %>% mutate(across(lu_shdi:field_pland, ~scale(.x)[,1]))

```

## Correlation between independent variables.

I selected a subset of variables that aren't strongly correlated and that represent potential drivers of ecosystem service patterns.

```{r correlation, echo = FALSE}

pairs.panels(wfs_lsm %>% 
               select(ann_agr_pland, prn_agr_pland, aglu_iji,
                      lu_shdi, aglu_frac_mn, margin_dens), 
             cex.cor = 1, cex.labels=1, cex=0.5, method = "pearson")

```

## Pollinator models

### Pollinator delta

```{r pollinator_delta}

pol.lm0 <- gam(delta_pollinators ~ margin_dens, data = wfs_lsm)
pol.gam0 <- gam(delta_pollinators ~ s(margin_dens), data = wfs_lsm)

anova(pol.lm0, pol.gam0)

pol.gam1 <- gam(delta_pollinators ~ s(ann_agr_pland) + s(aglu_iji) + s(prn_agr_pland) + s(aglu_frac_mn) + s(margin_dens) + s(lu_shdi), 
                data = wfs_lsm)

performance::compare_performance(pol.lm0, pol.gam0, pol.gam1)

summary(pol.gam1)

concurvity(pol.gam1)

```

```{r plot_pol}

polplot.df <- plot_fun(pol.gam1, 
                       c("ann_agr_pland", "aglu_iji", "prn_agr_pland", "aglu_frac_mn", "margin_dens"),
                       wfs_lsm)

ggplot(polplot.df[[1]], aes(x, delta_pollinators)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = polplot.df[[2]], 
             mapping = aes(color = margin_dens), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

```{r plot_pol0}

polplot0.df <- plot_fun(pol.gam0, 
                       "margin_dens",
                       wfs_lsm)

ggplot(polplot0.df[[1]], aes(x, delta_pollinators)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = polplot0.df[[2]], 
             mapping = aes(color = margin_dens), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

### Residual model from gam0

Trying to account for the effect of margins first

```{r pol_margin_resid}

wfs_lsm_resids <- wfs_lsm %>% cbind(pol.resid = residuals(pol.gam0))

polresid.gam1 <- gam(pol.resid ~ s(ann_agr_pland) + s(aglu_iji) + s(prn_agr_pland) + s(aglu_frac_mn) + s(lu_shdi), 
                data = wfs_lsm_resids)

summary(polresid.gam1)

concurvity(polresid.gam1)

```

```{r plot_pol_resid}

residpolplot.df <- plot_fun(polresid.gam1, 
                       c("ann_agr_pland", "aglu_iji", "prn_agr_pland", "aglu_frac_mn"),
                       wfs_lsm_resids,
                       "ann_agr_pland")

ggplot(residpolplot.df[[1]], aes(x, pol.resid)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = residpolplot.df[[2]], 
             mapping = aes(color = ann_agr_pland), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

## Seasonal water yield

### Seasonal water yield delta

```{r swy_delta}

swy.lm0 <- gam(delta_qb ~ margin_dens, data = wfs_lsm)
swy.gam0 <- gam(delta_qb ~ s(margin_dens), data = wfs_lsm)

anova(swy.lm0, swy.gam0)

swy.gam1 <- gam(delta_qb ~ s(ann_agr_pland) + s(aglu_iji) + s(prn_agr_pland) + s(aglu_frac_mn) + s(margin_dens) + s(lu_shdi), 
                data = wfs_lsm)

performance::compare_performance(swy.lm0, swy.gam0, swy.gam1)

summary(swy.gam1)

concurvity(swy.gam1)

```

```{r plot_swy}

swyplot.df <- plot_fun(swy.gam1, 
                       c("ann_agr_pland", "aglu_iji", "prn_agr_pland", "aglu_frac_mn", "margin_dens"),
                       wfs_lsm)

ggplot(swyplot.df[[1]], aes(x, delta_qb)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = swyplot.df[[2]], 
             mapping = aes(color = margin_dens), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

```{r plot_swy0}

swyplot0.df <- plot_fun(swy.gam0, 
                       "margin_dens",
                       wfs_lsm)

ggplot(swyplot0.df[[1]], aes(x, delta_qb)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = swyplot0.df[[2]], 
             mapping = aes(color = margin_dens), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

### Residual model from gam0

Trying to account for the effect of margins first

```{r swy_margin_resid}

wfs_lsm_resids <- wfs_lsm_resids %>% cbind(swy.resid = residuals(swy.gam0))

swyresid.gam1 <- gam(swy.resid ~ s(ann_agr_pland) + s(aglu_iji) + s(prn_agr_pland) + s(aglu_frac_mn) + s(lu_shdi), 
                data = wfs_lsm_resids)

summary(swyresid.gam1)

concurvity(swyresid.gam1)

```

```{r plot_swy_resid}

residswyplot.df <- plot_fun(swyresid.gam1, 
                       c("ann_agr_pland", "aglu_iji", "prn_agr_pland", "aglu_frac_mn"),
                       wfs_lsm_resids,
                       "ann_agr_pland")

ggplot(residswyplot.df[[1]], aes(x, swy.resid)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = residswyplot.df[[2]], 
             mapping = aes(color = ann_agr_pland), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

## Nutrient delivery ratio

### Nutrient delivery ratio delta

```{r ndr_delta}

ndr.lm0 <- gam(delta_n_export ~ margin_dens, data = wfs_lsm)
ndr.gam0 <- gam(delta_n_export ~ s(margin_dens), data = wfs_lsm)

anova(ndr.lm0, ndr.gam0)

ndr.gam1 <- gam(delta_n_export ~ s(ann_agr_pland) + s(aglu_iji) + s(prn_agr_pland) + s(aglu_frac_mn) + s(margin_dens) + s(lu_shdi), 
                data = wfs_lsm)

performance::compare_performance(ndr.lm0, ndr.gam0, ndr.gam1)

summary(ndr.gam1)

concurvity(ndr.gam1)

```

```{r plot_ndr}

ndrplot.df <- plot_fun(ndr.gam1, 
                       c("ann_agr_pland", "aglu_iji", "prn_agr_pland", "aglu_frac_mn", "margin_dens", "lu_shdi"),
                       wfs_lsm)

ggplot(ndrplot.df[[1]], aes(x, delta_n_export)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = ndrplot.df[[2]], 
             mapping = aes(color = margin_dens), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

```{r plot_ndr0}

ndrplot0.df <- plot_fun(ndr.gam0, 
                       "margin_dens",
                       wfs_lsm)

ggplot(ndrplot0.df[[1]], aes(x, delta_n_export)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = ndrplot0.df[[2]], 
             mapping = aes(color = margin_dens), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

### Residual model from gam0

Trying to account for the effect of margins first

```{r ndr_margin_resid}

wfs_lsm_resids <- wfs_lsm_resids %>% cbind(ndr.resid = residuals(ndr.gam0))

ndrresid.gam1 <- gam(ndr.resid ~ s(ann_agr_pland) + s(aglu_iji) + s(prn_agr_pland) + s(aglu_frac_mn) + s(lu_shdi), 
                data = wfs_lsm_resids)

summary(ndrresid.gam1)

concurvity(ndrresid.gam1)

```

```{r plot_ndr_resid}

residndrplot.df <- plot_fun(ndrresid.gam1, 
                       c("ann_agr_pland", "aglu_iji", "prn_agr_pland", "aglu_frac_mn", "lu_shdi"),
                       wfs_lsm_resids,
                       "ann_agr_pland")

ggplot(residndrplot.df[[1]], aes(x, ndr.resid)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = residndrplot.df[[2]], 
             mapping = aes(color = ann_agr_pland), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

## Sediment delivery ratio

### Sediment delivery ratio delta

```{r sdr_delta}

sdr.lm0 <- gam(delta_sed ~ margin_dens, data = wfs_lsm)
sdr.gam0 <- gam(delta_sed ~ s(margin_dens), data = wfs_lsm)

anova(sdr.lm0, sdr.gam0)

sdr.gam1 <- gam(delta_sed ~ s(ann_agr_pland) + s(aglu_iji) + s(prn_agr_pland) + s(aglu_frac_mn) + s(margin_dens) + s(lu_shdi), 
                data = wfs_lsm)

performance::compare_performance(sdr.lm0, sdr.gam0, sdr.gam1)

summary(sdr.gam1)

concurvity(sdr.gam1)

```

```{r plot_sdr}

sdrplot.df <- plot_fun(sdr.gam1, 
                       c("ann_agr_pland", "prn_agr_pland", "margin_dens"),
                       wfs_lsm)

ggplot(sdrplot.df[[1]], aes(x, delta_sed)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = sdrplot.df[[2]], 
             mapping = aes(color = margin_dens), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

```{r plot_sdr0}

sdrplot0.df <- plot_fun(sdr.gam0, 
                       "margin_dens",
                       wfs_lsm)

ggplot(sdrplot0.df[[1]], aes(x, delta_sed)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = sdrplot0.df[[2]], 
             mapping = aes(color = margin_dens), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

### Residual model from gam0

Trying to account for the effect of margins first

```{r sdr_margin_resid}

wfs_lsm_resids <- wfs_lsm_resids %>% cbind(sdr.resid = residuals(sdr.gam0))

sdrresid.gam1 <- gam(sdr.resid ~ s(ann_agr_pland) + s(aglu_iji) + s(prn_agr_pland) + s(aglu_frac_mn) + s(lu_shdi), 
                data = wfs_lsm_resids)

summary(sdrresid.gam1)

concurvity(sdrresid.gam1)

```

```{r plot_sdr_resid}

residsdrplot.df <- plot_fun(sdrresid.gam1, 
                       c("ann_agr_pland", "aglu_iji", "prn_agr_pland", "aglu_frac_mn"),
                       wfs_lsm_resids,
                       "ann_agr_pland")

ggplot(residsdrplot.df[[1]], aes(x, sdr.resid)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = residsdrplot.df[[2]], 
             mapping = aes(color = ann_agr_pland), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```