### 2024-12-27 landscapemodels_margin.Rmd

---
title: "R Notebook"
---

```{r setup, echo = FALSE}
library(tidyverse)
library(performance)
library(psych)
library(mgcv)
library(ggeffects)
library(piecewiseSEM)
source("R/utils-plot.r")

wfs <- readRDS("data/maxwfsresults_20241217_fix.rds")
lsm <- read.csv("data/lsm_reduced.csv") %>% 
  mutate(hu = gsub("hu", "", hu))

wfs_lsm <- wfs %>% left_join(lsm, by = "hu") %>%
  select(c(hu, huc2, name, margin.ha, lu_ta,
           pollinators_base:sed_export_base, 
           pollinators_max:sed_export_max, 
           delta_pollinators:delta_sed, 
           fields_max, field.ha, margin.ha, 
           lu_shdi:ntr_frac_mn)) %>%
  mutate(prn_ann_ratio = log(prn_agr_pland/ann_agr_pland),
         prn_pagr = 100*(prn_agr_pland/(ann_agr_pland+prn_agr_pland)),
         field_pland = field.ha/lu_ta,
         margin_dens = margin.ha/lu_ta,
         delta_pollinators.margin = delta_pollinators/margin.ha,
         delta_qb.margin = delta_qb/margin.ha,
         delta_n_export.margin = delta_n_export/margin.ha,
         delta_sed.margin = delta_sed/margin.ha)

scaled.data <- wfs_lsm %>% mutate(across(lu_shdi:field_pland, ~scale(.x)[,1]))

```

## Correlation between independent variables.

I selected a subset of variables that aren't strongly correlated and that represent potential drivers of ecosystem service patterns.

```{r correlation, echo = FALSE}

pairs.panels(wfs_lsm %>% 
               select(ann_agr_pland, prn_agr_pland, aglu_iji,
                      lu_shdi, aglu_frac_mn, margin_dens), 
             cex.cor = 1, cex.labels=1, cex=0.5, method = "pearson")

```

## Pollinator models

### Pollinator delta

```{r pollinator_delta}

pol.lm0 <- gam(delta_pollinators.margin ~ margin_dens, data = wfs_lsm)

summary(pol.lm0)

pol.gam1 <- gam(delta_pollinators.margin ~ s(ann_agr_pland) + s(aglu_iji) + s(prn_agr_pland) + s(aglu_frac_mn) + s(lu_shdi), 
                data = wfs_lsm)

performance::compare_performance(pol.lm0,pol.gam1)

summary(pol.gam1)

concurvity(pol.gam1)

```

```{r plot_pol}

polplot.df <- plot_fun(pol.gam1, 
                       c("ann_agr_pland", "aglu_iji", "prn_agr_pland", "aglu_frac_mn", "lu_shdi"),
                       wfs_lsm, "ann_agr_pland")

ggplot(polplot.df[[1]], aes(x, delta_pollinators.margin)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = polplot.df[[2]], 
             mapping = aes(color = ann_agr_pland), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

### Seasonal water yield delta

```{r swy_delta}

swy.lm0 <- gam(delta_qb.margin ~ margin_dens, data = wfs_lsm)

summary(swy.lm0)

swy.gam1 <- gam(delta_qb.margin ~ s(ann_agr_pland) + s(aglu_iji) + s(prn_agr_pland) + s(aglu_frac_mn) + s(lu_shdi), 
                data = wfs_lsm)

performance::compare_performance(swy.lm0,swy.gam1)

summary(swy.gam1)

concurvity(swy.gam1)

```

```{r plot_swy}

swyplot.df <- plot_fun(swy.gam1, 
                       c("ann_agr_pland", "aglu_iji", "prn_agr_pland", "aglu_frac_mn", "lu_shdi"),
                       wfs_lsm, "ann_agr_pland")

ggplot(swyplot.df[[1]], aes(x, delta_qb.margin)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = swyplot.df[[2]], 
             mapping = aes(color = ann_agr_pland), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

### Nutrient delivery delta

```{r ndr_delta}

ndr.lm0 <- gam(delta_n_export.margin ~ margin_dens, data = wfs_lsm)

summary(ndr.lm0)

ndr.gam1 <- gam(delta_n_export.margin ~ s(ann_agr_pland) + s(aglu_iji) + s(prn_agr_pland) + s(aglu_frac_mn) + s(lu_shdi), 
                data = wfs_lsm)

performance::compare_performance(ndr.lm0,ndr.gam1)

summary(ndr.gam1)

concurvity(ndr.gam1)

```

```{r plot_ndr}

ndrplot.df <- plot_fun(ndr.gam1, 
                       c("ann_agr_pland", "aglu_iji", "prn_agr_pland", "aglu_frac_mn", "lu_shdi"),
                       wfs_lsm, "ann_agr_pland")

ggplot(ndrplot.df[[1]], aes(x, delta_n_export.margin)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = ndrplot.df[[2]], 
             mapping = aes(color = ann_agr_pland), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

### Sediment delivery ratio delta

```{r sdr_delta}

sdr.lm0 <- gam(delta_sed.margin ~ margin_dens, data = wfs_lsm)

summary(sdr.lm0)

sdr.gam1 <- gam(delta_sed.margin ~ s(ann_agr_pland) + s(aglu_iji) + s(prn_agr_pland) + s(aglu_frac_mn) + s(lu_shdi), 
                data = wfs_lsm)

performance::compare_performance(sdr.lm0,sdr.gam1)

summary(sdr.gam1)

concurvity(sdr.gam1)

```

```{r plot_sdr}

sdrplot.df <- plot_fun(sdr.gam1, 
                       c("ann_agr_pland", "aglu_iji", "prn_agr_pland", "aglu_frac_mn", "lu_shdi"),
                       wfs_lsm, "ann_agr_pland")

ggplot(sdrplot.df[[1]], aes(x, delta_sed.margin)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = sdrplot.df[[2]], 
             mapping = aes(color = ann_agr_pland), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```