### 2025-01-15 landscapemodels_groups.Rmd
update the site with newest analysis

---
title: "Landscape analysis grouped by natural habitat level"
---

```{r setup, echo = FALSE, message=FALSE}
library(tidyverse)
library(performance)
library(psych)
library(mgcv)
library(ggeffects)
library(mgcViz)
source("R/utils-plot.r")

wfs <- readRDS("data/maxwfsresults_20241217_fix.rds")
lsm <- read.csv("data/lsm_reduced.csv") %>% 
  mutate(hu = gsub("hu", "", hu))

wfs_lsm <- wfs %>% left_join(lsm, by = "hu") %>%
  select(c(hu, huc2, name, margin.ha, lu_ta,
           pollinators_base:sed_export_base, 
           pollinators_max:sed_export_max, 
           delta_pollinators:delta_sed, 
           fields_max, field.ha, margin.ha, 
           lu_shdi:ntr_frac_mn)) %>%
  mutate(prn_ann_ratio = log(prn_agr_pland/ann_agr_pland),
         prn_pagr = 100*(prn_agr_pland/(ann_agr_pland+prn_agr_pland)),
         prn_ann_lograt = log(prn_agr_pland/ann_agr_pland),
         field_pland = field.ha/lu_ta,
         margin_dens = margin.ha/lu_ta,
         delta_pollinators.margin = delta_pollinators/margin_dens,
         delta_qb.margin = delta_qb/margin_dens,
         delta_n_export.margin = delta_n_export/margin_dens,
         delta_sed.margin = delta_sed/margin_dens) %>%
  mutate(habcat = cut(ntr_pland, breaks = c(-Inf,20,50,Inf),
                      labels = c("simple","moderate","natural")))

scaled.data <- wfs_lsm %>% mutate(across(lu_shdi:field_pland, ~scale(.x)[,1]))

wfs_sim <- wfs_lsm %>% filter(habcat == "simple")
wfs_mod <- wfs_lsm %>% filter(habcat == "moderate")
wfs_nat <- wfs_lsm %>% filter(habcat == "natural")

```

## Pollinator models

### Simple landscapes (<20% natural habitat)

#### Correlation between independent variables in simple landscapes.

```{r correlation_sim, echo = FALSE}

pairs.panels(wfs_sim %>% 
               select(ann_agr_pland, prn_agr_pland, aglu_iji,
                      lu_shdi, aglu_frac_mn, ntr_pland, prn_ann_lograt), 
             cex.cor = 1, cex.labels=1, cex=0.5, method = "pearson")

```

#### Pollinator delta for simple landscapes

```{r pollinator_delta_simple}

dpol.sim.gam <- gam(delta_pollinators.margin ~
                  s(ann_agr_pland) +
                  s(aglu_iji) +
                  s(prn_agr_pland) +
                  s(aglu_frac_mn) +
                  s(lu_shdi) + 
                  s(lu_ta), 
                data = wfs_sim)

summary(dpol.sim.gam)

concurvity(dpol.sim.gam, full = FALSE)

# gam.check(dpol.sim.gam)

```

Partial effects plots and partial residuals. If variables are highly correlated, variance in the partial residual plot can be much less than the actual variance: https://en.wikipedia.org/wiki/Partial_residual_plot

```{r part_plot_sim, echo=FALSE}

dpol.sim.viz <- getViz(dpol.sim.gam)

dpol.sim.vizplot <- plot(dpol.sim.viz, allTerms = T) + l_points() + l_fitLine(linetype = 3) + l_fitContour() + 
      l_ciLine(colour = 2) + l_ciBar()

print(dpol.sim.vizplot, pages = 1)

```

Predicted effects against actual points. Covariates of each plot(not on x axis) held at simple landscape averages.

```{r plot_pol_sim, echo=FALSE}

dpol.sim.pred <- plot_fun(dpol.sim.gam, 
                       c("ann_agr_pland", "aglu_iji", "prn_agr_pland",
                         "aglu_frac_mn", "lu_shdi", "lu_ta"),
                       wfs_sim, "ann_agr_pland")

ggplot(dpol.sim.pred[[1]], aes(x, delta_pollinators.margin)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = dpol.sim.pred[[2]], 
             mapping = aes(color = ann_agr_pland), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

#### Without perennial agriculture or Shannon diversity

These two variables increased concurvity when included with annual agriculture

```{r pollinator_delta_simple2}

dpol.sim.gam2 <- gam(delta_pollinators.margin ~
                  s(ann_agr_pland) +
                  s(aglu_iji) +
                  s(aglu_frac_mn) +
                  # s(lu_shdi) + 
                  s(lu_ta), 
                data = wfs_sim)

summary(dpol.sim.gam2)

concurvity(dpol.sim.gam2)

```

Partial effects plots and partial residuals

```{r part_plot_sim2, echo=FALSE}

dpol.sim.viz2 <- getViz(dpol.sim.gam2)

dpol.sim.vizplot2 <- plot(dpol.sim.viz2, allTerms = T) + l_points() + l_fitLine(linetype = 3) + l_fitContour() + 
      l_ciLine(colour = 2) + l_ciBar()

print(dpol.sim.vizplot2, pages = 1)

```

Predicted effects against actual points. Covariates of each plot(not on x axis) held at simple landscape averages.

```{r plot_pol_sim2, echo=FALSE}

dpol.sim.pred2 <- plot_fun(dpol.sim.gam2, 
                       c("ann_agr_pland", "aglu_iji",
                         "aglu_frac_mn", "lu_ta"),
                       wfs_sim, "ann_agr_pland")

ggplot(dpol.sim.pred2[[1]], aes(x, delta_pollinators.margin)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = dpol.sim.pred2[[2]], 
             mapping = aes(color = ann_agr_pland), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

#### A composite variable of log ratio of annual to perennial ag

The two types of ag become relative to each other so the connection to the original natural habitat gradient is lost. Therefore I added `ntr_pland` into the model.

```{r pollinator_delta_simple3}

dpol.sim.gam3 <- gam(delta_pollinators.margin ~
                  s(prn_ann_lograt) +
                  s(ntr_pland) +
                  s(aglu_iji) +
                  s(aglu_frac_mn) +
                  # s(lu_shdi) +
                  s(lu_ta), 
                data = wfs_sim)

summary(dpol.sim.gam3)

concurvity(dpol.sim.gam3)

```

Partial effects plots and partial residuals

```{r part_plot_sim3, echo=FALSE}

dpol.sim.viz3 <- getViz(dpol.sim.gam3)

dpol.sim.vizplot3 <- plot(dpol.sim.viz3, allTerms = T) + l_points() + l_fitLine(linetype = 3) + l_fitContour() + 
      l_ciLine(colour = 2) + l_ciBar()

print(dpol.sim.vizplot3, pages = 1)

```

Predicted effects against actual points. Covariates of each plot(not on x axis) held at simple landscape averages.

```{r plot_pol_sim3, echo=FALSE}

dpol.sim.pred3 <- plot_fun(dpol.sim.gam3, 
                       c("prn_ann_lograt", "ntr_pland", "aglu_iji",
                         "aglu_frac_mn", "lu_ta"),
                       wfs_sim, "prn_ann_lograt")

ggplot(dpol.sim.pred3[[1]], aes(x, delta_pollinators.margin)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = dpol.sim.pred3[[2]], 
             mapping = aes(color = prn_ann_lograt), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

### Pollinator delta for moderate landscapes (20-50% natural habitat)

#### Correlation between predictors

```{r correlation_mod}

pairs.panels(wfs_mod %>% 
               select(ann_agr_pland, prn_agr_pland, aglu_iji,
                      lu_shdi, aglu_frac_mn, ntr_pland, prn_ann_lograt), 
             cex.cor = 1, cex.labels=1, cex=0.5, method = "pearson")

```

Like in simple landscapes, strong concurvity.

```{r pollinator_delta_mod}

dpol.mod.gam <- gam(delta_pollinators.margin ~
                  s(ann_agr_pland) +
                  s(aglu_iji) +
                  s(prn_agr_pland) +
                  s(aglu_frac_mn) +
                  s(lu_shdi) + 
                  s(lu_ta), 
                data = wfs_mod)

summary(dpol.mod.gam)

concurvity(dpol.mod.gam, full=FALSE)

```

Partial effects plots and partial residuals

```{r part_plot_mod, echo=FALSE}

dpol.mod.viz <- getViz(dpol.mod.gam)

dpol.mod.vizplot <- plot(dpol.mod.viz, allTerms = T) + l_points() + l_fitLine(linetype = 3) + l_fitContour() + 
      l_ciLine(colour = 2) + l_ciBar()

print(dpol.mod.vizplot, pages = 1)

```

Predicted effects against actual points. Covariates of each plot(not on x axis) held at moderate landscape averages.

```{r plot_pol_mod, echo=FALSE}

dpol.mod.pred <- plot_fun(dpol.mod.gam, 
                       c("ann_agr_pland", "aglu_iji", "prn_agr_pland",
                         "aglu_frac_mn", "lu_shdi"),
                       wfs_mod, "ann_agr_pland")

ggplot(dpol.mod.pred[[1]], aes(x, delta_pollinators.margin)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = dpol.mod.pred[[2]], 
             mapping = aes(color = ann_agr_pland), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

#### Without perennial ag and landscape iji

These variables increased concurvity when included with annual agriculture

```{r pollinator_delta_mod2}

dpol.mod.gam2 <- gam(delta_pollinators.margin ~
                  s(ann_agr_pland) +
                  # s(aglu_iji) +
                  # s(prn_agr_pland) +
                  s(aglu_frac_mn) +
                  s(lu_shdi) +
                  s(lu_ta), 
                data = wfs_mod)

summary(dpol.mod.gam2)

concurvity(dpol.mod.gam2, full = TRUE)

```

Partial effects plots and partial residuals

```{r part_plot_mod2, echo=FALSE}

dpol.mod.viz2 <- getViz(dpol.mod.gam2)

dpol.mod.vizplot2 <- plot(dpol.mod.viz2, allTerms = T) + l_points() + l_fitLine(linetype = 3) + l_fitContour() + 
      l_ciLine(colour = 2) + l_ciBar()

print(dpol.mod.vizplot2, pages = 1)

```

Predicted effects against actual points. Covariates of each plot(not on x axis) held at moderate landscape averages.

```{r plot_pol_mod2, echo=FALSE}

dpol.mod.pred2 <- plot_fun(dpol.mod.gam2, 
                       c("ann_agr_pland", "lu_shdi",
                         "aglu_frac_mn", "lu_ta"),
                       wfs_mod, "ann_agr_pland")

ggplot(dpol.mod.pred2[[1]], aes(x, delta_pollinators.margin)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = dpol.mod.pred2[[2]], 
             mapping = aes(color = ann_agr_pland), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

#### A composite variable of log ratio of annual to perennial ag

Using the composite variable isn't as helpful now because the log ratio of perennial to annual ag is correlated with natural habitat, iji, and Shannon diversity.

```{r pollinator_delta_mod3}

dpol.mod.gam3 <- gam(delta_pollinators.margin ~
                  s(prn_ann_lograt) +
                  s(ntr_pland) +
                  # s(aglu_iji) +
                  s(aglu_frac_mn) +
                  # s(lu_shdi) +
                  s(lu_ta), 
                data = wfs_mod)

summary(dpol.mod.gam3)

concurvity(dpol.mod.gam3, full = TRUE)

```

Partial effects plots and partial residuals

```{r part_plot_mod3, echo=FALSE}

dpol.mod.viz3 <- getViz(dpol.mod.gam3)

dpol.mod.vizplot3 <- plot(dpol.mod.viz3, allTerms = T) + l_points() + l_fitLine(linetype = 3) + l_fitContour() + 
      l_ciLine(colour = 2) + l_ciBar()

print(dpol.mod.vizplot3, pages = 1)

```

Predicted effects against actual points. Covariates of each plot(not on x axis) held at moderate landscape averages.

```{r plot_pol_mod3, echo=FALSE}

dpol.mod.pred3 <- plot_fun(dpol.mod.gam3, 
                       c("prn_ann_lograt", "ntr_pland", 
                         "aglu_frac_mn", "lu_ta"),
                       wfs_mod, "prn_ann_lograt")

ggplot(dpol.mod.pred3[[1]], aes(x, delta_pollinators.margin)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = dpol.mod.pred3[[2]], 
             mapping = aes(color = prn_ann_lograt), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

### Pollinator delta for natural landscapes (>50% natural)

#### Correlation between predictors

```{r correlation_nat}

pairs.panels(wfs_nat %>% 
               select(ann_agr_pland, prn_agr_pland, aglu_iji,
                      lu_shdi, aglu_frac_mn, ntr_pland, prn_ann_lograt), 
             cex.cor = 1, cex.labels=1, cex=0.5, method = "pearson")

```

```{r pollinator_delta_nat}

dpol.nat.gam <- gam(delta_pollinators.margin ~
                  s(ann_agr_pland) +
                  s(aglu_iji) +
                  s(prn_agr_pland) +
                  s(aglu_frac_mn) +
                  s(lu_shdi) + 
                  s(lu_ta), 
                data = wfs_nat)

summary(dpol.nat.gam)

concurvity(dpol.nat.gam, full = FALSE)

```

Partial effects plots and partial residuals

```{r part_plot_nat, echo=FALSE}

dpol.nat.viz <- getViz(dpol.nat.gam)

dpol.nat.vizplot <- plot(dpol.nat.viz, allTerms = T) + l_points() + l_fitLine(linetype = 3) + l_fitContour() + 
      l_ciLine(colour = 2) + l_ciBar()

print(dpol.nat.vizplot, pages = 1)

```

Predicted effects against actual points. Covariates of each plot(not on x axis) held at natural landscape averages.

```{r plot_pol_nat, echo=FALSE}

dpol.nat.pred <- plot_fun(dpol.nat.gam, 
                       c("ann_agr_pland", "aglu_iji", "prn_agr_pland",
                         "aglu_frac_mn", "lu_shdi", "lu_ta"),
                       wfs_nat, "ann_agr_pland")

ggplot(dpol.nat.pred[[1]], aes(x, delta_pollinators.margin)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = dpol.nat.pred[[2]], 
             mapping = aes(color = ann_agr_pland), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

#### Without landscape iji

IJI increased concurvity when included with annual agriculture.

```{r pollinator_delta_nat2}

dpol.nat.gam2 <- gam(delta_pollinators.margin ~
                  s(ann_agr_pland) +
                  # s(aglu_iji) +
                  s(prn_agr_pland) +
                  s(aglu_frac_mn) +
                  s(lu_shdi) +
                  s(lu_ta), 
                data = wfs_nat)

summary(dpol.nat.gam2)

concurvity(dpol.nat.gam2, full = TRUE)

```

Partial effects plots and partial residuals

```{r part_plot_nat2, echo=FALSE}

dpol.nat.viz2 <- getViz(dpol.nat.gam2)

dpol.nat.vizplot2 <- plot(dpol.nat.viz2, allTerms = T) + l_points() + l_fitLine(linetype = 3) + l_fitContour() + 
      l_ciLine(colour = 2) + l_ciBar()

print(dpol.nat.vizplot2, pages = 1)

```

Predicted effects against actual points. Covariates of each plot(not on x axis) held at natural landscape averages.

```{r plot_pol_nat2, echo=FALSE}

dpol.nat.pred2 <- plot_fun(dpol.nat.gam2, 
                       c("ann_agr_pland", "lu_shdi", "prn_agr_pland",
                         "aglu_frac_mn", "lu_ta"),
                       wfs_nat, "ann_agr_pland")

ggplot(dpol.nat.pred2[[1]], aes(x, delta_pollinators.margin)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = dpol.nat.pred2[[2]], 
             mapping = aes(color = ann_agr_pland), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

#### A composite variable of log ratio of annual to perennial ag

Dropping IJI as before, but also Shannon diversity because it correlates with natural habitat

```{r pollinator_delta_nat3}

dpol.nat.gam3 <- gam(delta_pollinators.margin ~
                  s(prn_ann_lograt) +
                  s(ntr_pland) +
                  s(aglu_frac_mn) +
                  s(lu_ta), 
                data = wfs_nat)

summary(dpol.nat.gam3)

concurvity(dpol.nat.gam3, full = TRUE)

```

Partial effects plots and partial residuals

```{r part_plot_nat3, echo=FALSE}

dpol.nat.viz3 <- getViz(dpol.nat.gam3)

dpol.nat.vizplot3 <- plot(dpol.nat.viz3, allTerms = T) + l_points() + l_fitLine(linetype = 3) + l_fitContour() + 
      l_ciLine(colour = 2) + l_ciBar()

print(dpol.nat.vizplot3, pages = 1)

```

Predicted effects against actual points. Covariates of each plot(not on x axis) held at natural landscape averages.

```{r plot_pol_nat3, echo=FALSE}

dpol.nat.pred3 <- plot_fun(dpol.nat.gam3, 
                       c("prn_ann_lograt", "ntr_pland",
                         "aglu_frac_mn", "lu_ta"),
                       wfs_nat, "prn_ann_lograt")

ggplot(dpol.nat.pred3[[1]], aes(x, delta_pollinators.margin)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = dpol.nat.pred3[[2]], 
             mapping = aes(color = prn_ann_lograt), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  egg::theme_article()

```

## A complete dataset pollinator model attempt

#### Correlation between independent variables in all landscapes.

```{r correlation_all, echo = FALSE}

pairs.panels(wfs_lsm %>% 
               select(ann_agr_pland, prn_agr_pland, aglu_iji,
                      lu_shdi, aglu_frac_mn, ntr_pland, prn_ann_lograt), 
             cex.cor = 1, cex.labels=1, cex=0.5, method = "pearson")

```

```{r pol_complete}

dpol.all.gam <- gam(delta_pollinators.margin ~
                  te(prn_ann_lograt,ntr_pland) +
                  s(aglu_frac_mn) +
                  # s(aglu_iji) +
                  # s(lu_shdi) +
                  s(lu_ta), 
                data = wfs_lsm)

summary(dpol.all.gam)

concurvity(dpol.all.gam, full = TRUE)

```

Partial effects plots and partial residuals

```{r part_plot_all, echo=FALSE}
library(rgl)

dpol.all.viz <- getViz(dpol.all.gam)

dpol.all.vizplot <- plot(dpol.all.viz, allTerms = T) + l_points() + l_fitLine(linetype = 3) + l_fitContour() + 
      l_ciLine(colour = 2) + l_ciBar()

print(dpol.all.vizplot, pages = 1)

plotRGL(sm(dpol.all.viz, 1), fix = c("z" = 0), residuals = TRUE)
```