### 2025-02-17 landscapemetrics_quantile.Rmd
update analysis

---
title: "Quantile regression"
---

Quantile regression looks at relationships at different quantiles of response.

```{r setup, echo = FALSE, message=FALSE}
library(tidyverse)
library(qgam)
library(mgcViz)
library(ggfortify)

wfs_lsm <- readRDS("data/wfs_lsm.rds")

```

```{r qgam_fit}
fit <- mqgamV(delta_pollinators.margin ~
                  te(prn_ann_lograt,ntr_pland) +
                  s(aglu_frac_mn) +
                  s(aglu_iji) +
                  s(lu_shdi),
                  # s(lu_ta), 
                data = wfs_lsm,
              qu = c(0.2,0.5,0.8))

```

```{r qgam_vis}
print(plot(fit, allTerms = TRUE), pages = 1)
```

Okay that looks similar to what I've already done

```{r margin_relations}

wfs_low <- wfs_lsm %>% filter(ntr_pland < 8.25 & ntr_pland > 7.75)

pairs.panels(wfs_lsm %>% 
               select(delta_pollinators.margin, delta_qb.margin, 
                      delta_n_export.margin, delta_sed.margin), 
             cex.cor = 1, cex.labels=1, cex=0.5, method = "pearson")

```

```{r es_relations}

pairs.panels(wfs_lsm %>% 
               select(pollinators_base, qb_base, 
                      n_export_base, sed_export_base), 
             cex.cor = 1, cex.labels=1, cex=0.5, method = "pearson")

```

```{r quant_cut}

wfs_grp <- wfs_lsm %>%
  mutate(habcat = cut(ntr_pland, breaks = c(-Inf,20,50,Inf),
                      labels = c("simple","moderate","natural")))
```

```{r grp_pca_smp}

wfs_pca_all <- prcomp(wfs_grp %>%
                        select(ntr_pland, ann_agr_pland, prn_agr_pland, 
                               aglu_iji, aglu_frac_mn, lu_shdi),
                      scale. = TRUE, center = TRUE)

summary(wfs_pca_all)

```

```{r grp_pca_mod}

ggbiplot::ggbiplot(wfs_pca_all, groups = wfs_grp$habcat)
ggbiplot::ggbiplot(wfs_pca_all, groups = wfs_grp$habcat, choices = c(2,3))

```

```{r grp_pca_smp2}

wfs_pca_all2 <- prcomp(wfs_grp %>%
                        select(ntr_pland, ann_agr_pland, prn_agr_pland, 
                               aglu_iji, aglu_frac_mn),
                      scale. = TRUE, center = TRUE)

summary(wfs_pca_all2)

```

```{r grp_pca_mod2}

ggbiplot::ggbiplot(wfs_pca_all2, groups = wfs_grp$habcat)
ggbiplot::ggbiplot(wfs_pca_all2, groups = wfs_grp$habcat, choices = c(2,3))

```

```{r create_pc_var}

lsm_pc <- wfs_grp %>% cbind(predict(wfs_pca_all))

ggplot(lsm_pc, aes(PC1, pollinators_base, colour = habcat)) + geom_point()

ggplot(lsm_pc, aes(PC3, sed_export_base, colour = habcat)) + geom_point()

```