### 2025-02-26 landscapemetrics_onefield.Rmd
add one field data and documentation

---
title: "One Field"
output: html_document
date: "2025-02-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Add one field

```{r load}
library(tidyverse)
library(psych)
library(egg)

wfs_lsm <- readRDS("data/wfs_lsm.rds") 
of_lsm <- readRDS("data/of_lsm.rds")
of_lsm_margin.med <- readRDS("data/of_lsm_marginmed.rds")

```

```{r recombine_data}
wfs_lsm_pc <- readRDS("data/wfs_lsm_pc.rds")

of_lsm.med <- of_lsm %>% 
  group_by(hu) %>%
  summarize(
    across(delta_pollinators:delta_sed, 
           list(mean = mean, sd = sd, median = median),
           .names = "{.col}_of.{.fn}")
  )

wfs_lsm_of <- wfs_lsm_pc %>% 
  left_join(
    of_lsm_margin.med %>%
      select(hu, 
             delta_pollinators_of.margin.median,
             delta_qb_of.margin.median,
             delta_n_export_of.margin.median,
             delta_sed_of.margin.median
             )
  ) %>% 
  left_join(
    of_lsm.med %>%
      select(hu, 
             delta_pollinators_of.median,
             delta_qb_of.median,
             delta_n_export_of.median,
             delta_sed_of.median
             )
  )

pairs.panels(
  wfs_lsm_of %>% 
    select(
      delta_pollinators.margin,
      delta_qb.margin,
      delta_n_export.margin,
      delta_sed.margin,
      delta_pollinators_of.margin.median,
      delta_qb_of.margin.median,
      delta_n_export_of.margin.median,
      delta_sed_of.margin.median
    )
)

```

```{r pol}

pol_long <- wfs_lsm_of %>% 
  pivot_longer(cols= c(delta_pollinators.margin, delta_pollinators_of.margin.median),
               names_to = "treatment",
               names_prefix = "delta_pollinators.")

pol_long.ws <- wfs_lsm_of %>% 
  pivot_longer(cols= c(delta_pollinators, delta_pollinators_of.median),
               names_to = "treatment", values_to = "pollinators",
               names_prefix = "delta_pollinators.") %>%
  mutate(pollinators_ws = pollinators * lu_ta)

ggplot(pol_long, aes(ntr_pland, value, color = treatment)) + geom_point()
ggplot(pol_long.ws, aes(PC1, pollinators_ws, color = treatment)) + geom_point()

ggplot(wfs_lsm_of, aes(delta_pollinators.margin,delta_pollinators_of.margin.median)) + geom_point()

ggplot(wfs_lsm_of, aes(ntr_pland,delta_pollinators)) + geom_point()
ggplot(wfs_lsm_of, aes(ntr_pland,delta_pollinators_of.median)) + geom_point()

ggplot(wfs_lsm_of, aes(PC1,delta_pollinators)) + geom_point()
ggplot(wfs_lsm_of, aes(ntr_pland,delta_n_export)) + geom_point()

ggplot(wfs_lsm_of, aes(aglu_frac_mn,delta_pollinators.margin, color = habcat)) + geom_point()
ggplot(wfs_lsm_of, aes(PC1,delta_pollinators.margin)) + geom_point()
ggplot(wfs_lsm_of, aes(PC1,delta_pollinators_of.margin.median)) + geom_point()
```

```{r n_exp}

nexp_long <- wfs_lsm_of %>% 
  pivot_longer(cols= c(delta_n_export.margin, delta_n_export_of.margin.median),
               names_to = "treatment",
               names_prefix = "delta_n_export.")

ggplot(nexp_long, aes(PC1, value, color = treatment)) + geom_point()

```

```{r panel_base}

onefield_long <- wfs_lsm_of %>% 
  pivot_longer(cols= c(pollinators_base,
                       n_export_base,
                       qb_base,
                       sed_export_base),
               names_to = "Ecosystem service",
               values_to = "ES value")

ggplot(onefield_long, aes(PC2, `ES value`)) + geom_point() + 
  facet_wrap(~`Ecosystem service`, scales = "free_y")

```

```{r panel_onefield}

onefield_long <- wfs_lsm_of %>% 
  pivot_longer(cols= c(delta_pollinators_of.median,
                       delta_n_export_of.median,
                       delta_qb_of.median,
                       delta_sed_of.median),
               names_to = "Ecosystem service",
               values_to = "ES value",
               names_prefix = "delta_") %>%
  mutate(`ES value` = `ES value`*lu_ta)

ggplot(onefield_long, aes(PC2, `ES value`)) + geom_point() + 
  facet_wrap(~`Ecosystem service`, scales = "free_y")

```

```{r panel_max}
max_long <- wfs_lsm_of %>% 
  pivot_longer(cols= c(delta_pollinators,
                       delta_n_export,
                       delta_qb,
                       delta_sed),
               names_to = "Ecosystem service",
               values_to = "ES value",
               names_prefix = "delta_") %>%
  mutate(`ES value` = `ES value`*lu_ta)%>%
  mutate(
    `Ecosystem service` = fct_recode(
      `Ecosystem service`,
      "Pollinators" = "pollinators",
      "N export reduced" = "n_export",
      "Water recharge" = "qb",
      "Sediment reduced" = "sed"
    )
  )

ggplot(max_long, aes(PC1, `ES value`, color = "Treatment")) + geom_point(alpha = 0.5) + 
  facet_wrap(~`Ecosystem service`, scales = "free_y") + 
  theme_article() +
  theme(strip.text = element_text(size = 12))

```

```{r panel_onefield_margin}

onefield_long.margin <- wfs_lsm_of %>% 
  pivot_longer(cols= c(delta_pollinators_of.margin.median,
                       delta_n_export_of.margin.median,
                       delta_qb_of.margin.median,
                       delta_sed_of.margin.median),
               names_to = "Ecosystem service",
               values_to = "ES value",
               names_prefix = "delta_") %>%
  # filter(!(`Ecosystem service` == "sed_of.margin.median" & `ES value` > 60)) %>%
  mutate(
    `Ecosystem service` = fct_recode(
      `Ecosystem service`,
      "Pollinators" = "pollinators_of.margin.median",
      "N export reduced" = "n_export_of.margin.median",
      "Water recharge" = "qb_of.margin.median",
      "Sediment reduced" = "sed_of.margin.median"
    )
  )

ggplot(onefield_long.margin, aes(PC1, `ES value`, color = habcat)) + 
  geom_point() + 
  facet_wrap(~`Ecosystem service`, scales = "free_y") +
  theme_article()

ggplot(onefield_long.margin, aes(PC2, `ES value`, color = habcat)) + geom_point() + 
  facet_wrap(~`Ecosystem service`, scales = "free_y") +
  theme_article()

```

```{r panel_max_margin}

max_long.margin <- wfs_lsm_of %>% 
  pivot_longer(cols= c(delta_pollinators.margin,
                       delta_n_export.margin,
                       delta_qb.margin,
                       delta_sed.margin),
               names_to = "Ecosystem service",
               values_to = "ES value",
               names_prefix = "delta_") %>%
  mutate(
    `Ecosystem service` = fct_recode(
      `Ecosystem service`,
      "Pollinators" = "pollinators.margin",
      "N export reduced" = "n_export.margin",
      "Water recharge" = "qb.margin",
      "Sediment reduced" = "sed.margin"
    )
  )

ggplot(max_long.margin, aes(PC1, `ES value`, color = habcat)) +
  geom_point() + 
  facet_wrap(~`Ecosystem service`, scales = "free_y") +
  theme_article() +
  theme(strip.text = element_text(size = 12))

ggplot(max_long.margin, aes(PC2, `ES value`, color = habcat)) + 
  geom_point() + 
  facet_wrap(~`Ecosystem service`, scales = "free_y") +
  theme_article() +
  theme(strip.text = element_text(size = 12))

```

```{r both_types_margin}
both_long.margin <- bind_rows(
  onefield_long.margin %>% mutate(Treatment = "One field"),
  max_long.margin %>% mutate(Treatment = "Max fields")
)

ggplot(both_long.margin, aes(PC1, `ES value`, color = Treatment)) +
  geom_point(alpha = 0.5) + 
  facet_wrap(~`Ecosystem service`, scales = "free_y") +
  theme_article() +
  theme(strip.text = element_text(size = 12))

ggplot(both_long.margin, aes(PC2, `ES value`, color = Treatment)) +
  geom_point(alpha = 0.5) + 
  facet_wrap(~`Ecosystem service`, scales = "free_y") +
  theme_article() +
  theme(strip.text = element_text(size = 12))

```

```{r es_pca}

es_base_pca <- prcomp(wfs_lsm_of %>%
                        select(pollinators_base:sed_export_base),
                      scale. = TRUE, center = TRUE)

es_base_pca

summary(es_base_pca)

ggbiplot::ggbiplot(es_base_pca, groups = wfs_lsm_of$habcat) +
  egg::theme_article()

ggbiplot::ggbiplot(es_base_pca, groups = wfs_lsm_of$habcat, choices = c(2,3)) +
  egg::theme_article()
```

```{r es_margin_pca}

es_max.margin_pca <- prcomp(wfs_lsm_of %>%
                        select(delta_pollinators.margin:delta_sed.margin),
                      scale. = TRUE, center = TRUE)

es_max.margin_pca

summary(es_max.margin_pca)

ggbiplot::ggbiplot(es_max.margin_pca, groups = wfs_lsm_of$habcat) +
  egg::theme_article()

ggbiplot::ggbiplot(es_max.margin_pca, groups = wfs_lsm_of$habcat, choices = c(2,3)) +
  egg::theme_article()
```

```{r model}
library(brms)

both_wide.margin <- both_long.margin %>%
  pivot_wider(id_cols = c(hu, Treatment),values_from = `ES value`, names_from = `Ecosystem service`) %>% left_join(wfs_lsm_pc, by = "hu") %>%
  rename(
      pollinators = "Pollinators",
      N_exp_red = "N export reduced",
      water_rech = "Water recharge",
      sed_exp_red = "Sediment reduced"
  )

allmod <- brm(mvbind(pollinators, N_exp_red, water_rech, sed_exp_red) ~ Treatment * PC1 * PC2 + (1|p|hu), data = both_wide.margin)

allmod2 <- brm(mvbind(pollinators, N_exp_red, water_rech, sed_exp_red) ~ Treatment + t2(PC1, PC2, by = Treatment) + (1|p|hu), data = both_wide.margin)

```