### 2025-03-05 data-raw/examine_sim_results.Rmd
update site

---
title: "examine simulation results"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

##

```{r load}
wfs_lsm <- readRDS("data/wfs_lsm.rds")
hu_meta <- readRDS("data/hu_meta.rds")

hulandstats <- readRDS("data/hulandstats.rds")

wfs_meta <- wfs_lsm %>% 
  left_join(hu_meta %>% sf::st_drop_geometry(), 
            by = c("hu" = "huc12", "huc2", "name")) %>%
  left_join(hulandstats)

```

## Plots

```{r wetland}
ggplot(wfs_meta, aes(Wetland, delta_sed.margin)) + 
  geom_point()
```

```{r erosivity}
ggplot(wfs_meta, aes(erosivity_mean, delta_sed.margin)) + 
  geom_point()
```

```{r erodibility}
ggplot(wfs_meta, aes(erodibility_mean, delta_sed.margin)) + 
  geom_point()
```

```{r erodibility_2}
library(plotly)
fig <- plot_ly(wfs_meta, x = ~erodibility_mean, y = ~erosivity_mean, z = ~delta_sed.margin, color = ~delta_sed.margin, text = ~hu)

fig

```

```{r sediment_change}
ggplot(wfs_meta, aes(sed_export_base, sed_export_max, color = delta_sed.margin)) + 
  geom_point() + geom_abline(intercept = 0, slope = 1) + scale_color_continuous(type="viridis")

```

```{r nutrient_change}
ggplot(wfs_meta, aes(n_export_base, n_export_max, color = delta_n_export.margin)) + 
  geom_point() + geom_abline(intercept = 0, slope = 1) + scale_color_continuous(type="viridis")

```

```{r pollinator_change}
ggplot(wfs_meta, aes(pollinators_base, pollinators_max, color = delta_pollinators.margin)) + 
  geom_point() + geom_abline(intercept = 0, slope = 1) + scale_color_continuous(type="viridis")

```

```{r qb_change}
ggplot(wfs_meta, aes(qb_base, qb_max, color = delta_qb.margin)) + 
  geom_point() + geom_abline(intercept = 0, slope = 1) + scale_color_continuous(type="viridis")

```