### 2025-03-11 examine_sim_results.Rmd

---
title: "examine simulation results"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

##

```{r load, include = FALSE}
library(tidyverse)
# wfs_lsm <- readRDS("data/wfs_lsm.rds")
# hu_meta <- readRDS("data/hu_meta.rds")
# 
# hulandstats <- readRDS("data/hulandstats.rds")
# 
# 
# wfs_meta <- wfs_lsm %>% 
#   left_join(hu_meta %>% sf::st_drop_geometry(), 
#             by = c("hu" = "huc12", "huc2", "name")) %>%
#   left_join(hulandstats)

wfs_meta <- readRDS("data/allresults_20250308.rds")

```

```{r visualize_base}

base_df <- wfs_meta %>% 
  select(c(hu, ntr_pland, pollinators_base:sed_export_base)) %>%
  pivot_longer(cols = pollinators_base:sed_export_base,
               names_to = "Ecosystem_service", values_to = "Ecosystem_value")

ggplot(base_df, aes(ntr_pland, Ecosystem_value)) +
  geom_point() +
  facet_wrap(~Ecosystem_service, scales = "free_y") + egg::theme_article()

```

```{r visualize_delta}

delta_df <- wfs_meta %>% 
  select(c(hu, ntr_pland, delta_pollinators:delta_sed)) %>%
  pivot_longer(cols = delta_pollinators:delta_sed,
               names_to = "Ecosystem_service", values_to = "ES_value_change")

ggplot(delta_df, aes(ntr_pland, ES_value_change)) +
  geom_point() +
  facet_wrap(~Ecosystem_service, scales = "free_y") + egg::theme_article()

```

```{r visualize_delta_margin}

delta_df <- wfs_meta %>% 
  select(c(hu, ntr_pland, delta_pollinators_max.margin:delta_sed_max.margin)) %>%
  pivot_longer(cols = delta_pollinators_max.margin:delta_sed_max.margin,
               names_to = "Ecosystem_service", values_to = "ES_value_change_ha_margin")

ggplot(delta_df, aes(ntr_pland, ES_value_change_ha_margin)) +
  geom_point() +
  facet_wrap(~Ecosystem_service, scales = "free_y") + egg::theme_article()

```

```{r visualize_delta_margin_onefield}

delta_df <- wfs_meta %>% 
  select(c(hu, ntr_pland,
           delta_pollinators_of.margin.median, 
           delta_n_export_of.margin.median,
           delta_qb_of.margin.median,
           delta_sed_of.margin.median)) %>%
  pivot_longer(cols = delta_pollinators_of.margin.median:delta_sed_of.margin.median,
               names_to = "Ecosystem_service", 
               values_to = "One_field_change_ha_margin")

ggplot(delta_df, aes(ntr_pland, One_field_change_ha_margin)) +
  geom_point() +
  facet_wrap(~Ecosystem_service, scales = "free_y") + egg::theme_article()

```

## Plots

```{r wetland}
ggplot(wfs_meta, aes(Wetland, delta_sed_max.margin)) + 
  geom_point()
```

```{r erosivity}
ggplot(wfs_meta, aes(erosivity_mean, delta_sed_max.margin)) + 
  geom_point()
```

```{r erodibility}
ggplot(wfs_meta, aes(erodibility_mean, delta_sed_max.margin)) + 
  geom_point()
```

```{r erodibility_2}
library(plotly)
fig <- plot_ly(wfs_meta, x = ~erodibility_mean, y = ~erosivity_mean, z = ~delta_sed_max.margin, color = ~delta_sed_max.margin, text = ~hu)

fig

```

```{r sediment_change}
ggplot(wfs_meta, aes(sed_export_base, sed_export_max, color = delta_sed_max.margin)) + 
  geom_point() + geom_abline(intercept = 0, slope = 1) + scale_color_continuous(type="viridis")

```

```{r nutrient_change}
ggplot(wfs_meta, aes(n_export_base, n_export_max, color = delta_n_export_max.margin)) + 
  geom_point() + geom_abline(intercept = 0, slope = 1) + scale_color_continuous(type="viridis")

```

```{r pollinator_change}
ggplot(wfs_meta, aes(pollinators_base, pollinators_max, color = delta_pollinators_max.margin)) + 
  geom_point() + geom_abline(intercept = 0, slope = 1) + scale_color_continuous(type="viridis")

```

```{r qb_change}
ggplot(wfs_meta, aes(qb_base, qb_max, color = delta_qb_max.margin)) + 
  geom_point() + geom_abline(intercept = 0, slope = 1) + scale_color_continuous(type="viridis")

```