### 2025-03-11 old/landscapemodels.Rmd

---
title: "Landscape models"
---

```{r setup, echo = FALSE}
library(tidyverse)
library(performance)
library(psych)
source("R/utils-plot.r")

wfs <- readRDS("data/maxwfsresults_20241217_fix.rds")
lsm <- read.csv("data/lsm_reduced.csv") %>% 
  mutate(hu = gsub("hu", "", hu))

wfs_lsm <- wfs %>% left_join(lsm, by = "hu") %>%
  select(c(hu, huc2, name, margin.ha, lu_ta,
           pollinators_base:sed_export_base, 
           pollinators_max:sed_export_max, 
           delta_pollinators:delta_sed, 
           fields_max, field.ha, margin.ha, 
           lu_shdi:ntr_frac_mn)) %>%
  mutate(prn_ann_ratio = log(prn_agr_pland/ann_agr_pland),
         prn_pagr = 100*(prn_agr_pland/(ann_agr_pland+prn_agr_pland)),
         field_pland = field.ha/lu_ta,
         margin_dens = margin.ha/lu_ta)

scaled.data <- wfs_lsm %>% mutate(across(lu_shdi:field_pland, ~scale(.x)[,1]))

```

## Correlation between independent variables.

I selected a subset of variables that aren't strongly correlated and that represent potential drivers of ecosystem service patterns.

```{r correlation, echo = FALSE}

pairs.panels(wfs_lsm %>% 
               select(ann_agr_pland, prn_agr_pland, aglu_iji,
                      lu_shdi, aglu_frac_mn, margin_dens), 
             cex.cor = 1, cex.labels=1, cex=0.5)

```

## Pollinator models

### Pollinator delta

```{r pollinator_delta}

pol.dmod0 <- lm(delta_pollinators ~ margin_dens, data = scaled.data)
pol.dmod1 <- lm(delta_pollinators ~ ann_agr_pland + aglu_iji + prn_agr_pland + aglu_frac_mn + margin_dens + lu_shdi, data = scaled.data)
pol.dmod2 <- lm(delta_pollinators ~ ann_agr_pland + I(ann_agr_pland^2) + aglu_iji + prn_agr_pland + aglu_frac_mn + margin_dens + lu_shdi, data = scaled.data)

performance::compare_performance(pol.dmod0, pol.dmod1, pol.dmod2)

summary(pol.dmod2)

```

```{r pollinator_plots, echo = FALSE}

pol.vars <- c("margin_dens", "ann_agr_pland", "aglu_frac_mn", "aglu_iji","prn_agr_pland")

pol.preds <- plot_fun(pol.dmod2, pol.vars, scaled.data)

ggplot(pol.preds[[1]], aes(x, delta_pollinators)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = pol.preds[[2]], 
             mapping = aes(color = margin_dens), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  xlab("z-score") +
  egg::theme_article()

```

## Seasonal water yield

### Seasonal water yield delta

```{r swy_delta}

swy.dmod0 <- lm(delta_qb ~ margin_dens, data = scaled.data)
swy.dmod1 <- lm(delta_qb ~ ann_agr_pland + aglu_iji + prn_agr_pland + aglu_frac_mn + margin_dens + lu_shdi, data = scaled.data)
swy.dmod2 <- lm(delta_qb ~ ann_agr_pland + I(ann_agr_pland^2) + aglu_iji + prn_agr_pland + aglu_frac_mn + margin_dens + lu_shdi, data = scaled.data)

performance::compare_performance(swy.dmod0, swy.dmod1, swy.dmod2)

summary(swy.dmod2)

```

```{r swy_plots, echo = FALSE}

swy.vars <- c("margin_dens", "ann_agr_pland", "prn_agr_pland", "aglu_iji", "aglu_frac_mn")

swy.preds <- plot_fun(swy.dmod2, swy.vars, scaled.data)

ggplot(swy.preds[[1]], aes(x, delta_qb)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = swy.preds[[2]], 
             mapping = aes(color = margin_dens), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  xlab("z-score") +
  egg::theme_article()

```

## Nutrient delivery ratio

### Nutrient delivery ratio delta

```{r ndr_delta}

ndr.dmod0 <- lm(delta_n_export ~ margin_dens, data = scaled.data)
ndr.dmod1 <- lm(delta_n_export ~ ann_agr_pland + aglu_iji + prn_agr_pland + aglu_frac_mn + margin_dens + lu_shdi, data = scaled.data)
ndr.dmod2 <- lm(delta_n_export ~ ann_agr_pland + I(ann_agr_pland^2) + aglu_iji + prn_agr_pland + aglu_frac_mn + margin_dens + lu_shdi, data = scaled.data)

performance::compare_performance(ndr.dmod0, ndr.dmod1, ndr.dmod2)

summary(ndr.dmod2)

```

```{r ndr_plots, echo = FALSE}

ndr.vars <- c("margin_dens", "ann_agr_pland", "prn_agr_pland", 
              "aglu_iji", "aglu_frac_mn", "lu_shdi")

ndr.preds <- plot_fun(ndr.dmod2, ndr.vars, scaled.data)

ggplot(ndr.preds[[1]], aes(x, delta_n_export)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = ndr.preds[[2]], 
             mapping = aes(color = margin_dens), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  xlab("z-score") +
  egg::theme_article()

```

## Sediment delivery ratio

### Sediment delivery ratio delta

```{r sdr_delta}

sdr.dmod0 <- lm(delta_sed ~ margin_dens, data = scaled.data)
sdr.dmod1 <- lm(delta_sed ~ ann_agr_pland + aglu_iji + prn_agr_pland + aglu_frac_mn + margin_dens + lu_shdi, data = scaled.data)
sdr.dmod2 <- lm(delta_sed ~ ann_agr_pland + I(ann_agr_pland^2) + aglu_iji + prn_agr_pland + aglu_frac_mn + margin_dens + lu_shdi, data = scaled.data)

performance::compare_performance(sdr.dmod0, sdr.dmod1, sdr.dmod2)

summary(sdr.dmod2)

```

```{r sdr_plots, echo = FALSE}

sdr.vars <- c("margin_dens", "ann_agr_pland", "prn_agr_pland", "lu_shdi")

sdr.preds <- plot_fun(sdr.dmod2, sdr.vars, scaled.data)

ggplot(sdr.preds[[1]], aes(x, delta_sed)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = sdr.preds[[2]], 
             mapping = aes(color = margin_dens), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  xlab("z-score") +
  egg::theme_article()

```

## Residuals-based models

Since margin density is a major effect, I regress the other predictors against the residuals of univariate models with only margin density

### Pollinator model residuals

```{r pollinators_resid, echo = FALSE}

pol.resid <- data.frame(pol_resid = pol.dmod0$residuals, 
                        scaled.data %>% 
                          select(ann_agr_pland, prn_agr_pland, aglu_iji, 
                                 aglu_frac_mn, lu_shdi))

pol.resid.mod <- lm(pol_resid ~ ann_agr_pland + I(ann_agr_pland^2) + prn_agr_pland + aglu_iji + aglu_frac_mn + lu_shdi, data = pol.resid)

summary(pol.resid.mod)

pol.resid.vars <- c("ann_agr_pland", "prn_agr_pland", "aglu_frac_mn", "aglu_iji", "lu_shdi")

pol.resid.preds <- plot_fun(pol.resid.mod, pol.resid.vars, pol.resid, "ann_agr_pland")

ggplot(pol.resid.preds[[1]], aes(x, pol_resid)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = pol.resid.preds[[2]], 
             mapping = aes(color = ann_agr_pland), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  xlab("z-score") +
  egg::theme_article()

```

### Seasonal water yield residuals model

```{r swy_resid, echo = FALSE}

swy.resid <- data.frame(swy_resid = swy.dmod0$residuals, 
                        scaled.data %>% 
                          select(ann_agr_pland, prn_agr_pland, aglu_iji, 
                                 aglu_frac_mn, lu_shdi))

swy.resid.mod <- lm(swy_resid ~ ann_agr_pland + I(ann_agr_pland^2) + prn_agr_pland + aglu_iji + aglu_frac_mn + lu_shdi, data = swy.resid)

summary(swy.resid.mod)

swy.resid.vars <- c("ann_agr_pland", "prn_agr_pland", "aglu_frac_mn", "aglu_iji", "lu_shdi")

swy.resid.preds <- plot_fun(swy.resid.mod, swy.resid.vars, swy.resid, "ann_agr_pland")

ggplot(swy.resid.preds[[1]], aes(x, swy_resid)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = swy.resid.preds[[2]], 
             mapping = aes(color = ann_agr_pland), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  xlab("z-score") +
  egg::theme_article()

```

### Nutrient delivery residuals model

```{r ndr_resid, echo = FALSE}

ndr.resid <- data.frame(ndr_resid = ndr.dmod0$residuals, 
                        scaled.data %>% 
                          select(ann_agr_pland, prn_agr_pland, aglu_iji, 
                                 aglu_frac_mn, lu_shdi))

ndr.resid.mod <- lm(ndr_resid ~ ann_agr_pland + I(ann_agr_pland^2) + prn_agr_pland + aglu_iji + aglu_frac_mn + lu_shdi, data = ndr.resid)

summary(ndr.resid.mod)

ndr.resid.vars <- c("ann_agr_pland", "prn_agr_pland", "aglu_frac_mn", "aglu_iji", "lu_shdi")

ndr.resid.preds <- plot_fun(ndr.resid.mod, ndr.resid.vars, ndr.resid, "ann_agr_pland")

ggplot(ndr.resid.preds[[1]], aes(x, ndr_resid)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = ndr.resid.preds[[2]], 
             mapping = aes(color = ann_agr_pland), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  xlab("z-score") +
  egg::theme_article()

```

### Sediment delivery residuals model

```{r sdr_resid, echo = FALSE}

sdr.resid <- data.frame(sdr_resid = sdr.dmod0$residuals, 
                        scaled.data %>% 
                          select(ann_agr_pland, prn_agr_pland, aglu_iji, 
                                 aglu_frac_mn, lu_shdi))

sdr.resid.mod <- lm(sdr_resid ~ ann_agr_pland + I(ann_agr_pland^2) + prn_agr_pland + aglu_iji + aglu_frac_mn + lu_shdi, data = sdr.resid)

summary(sdr.resid.mod)

sdr.resid.vars <- c("ann_agr_pland", "prn_agr_pland", "aglu_frac_mn", "aglu_iji", "lu_shdi")

sdr.resid.preds <- plot_fun(sdr.resid.mod, sdr.resid.vars, sdr.resid, "ann_agr_pland")

ggplot(sdr.resid.preds[[1]], aes(x, sdr_resid)) +  
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
  geom_point(data = sdr.resid.preds[[2]], 
             mapping = aes(color = ann_agr_pland), 
             size = 0.6) + scale_color_continuous(type = "viridis") +
  facet_wrap(~pred.var, scales = "free_x") +
  xlab("z-score") +
  egg::theme_article()

```