### 2025-03-26 Paper_analysis.Rmd
update site

---
title: "Paper analysis"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r load}
library(tidyverse)
load("data/ES-RF2.RDA")

```

## Patterns with composition

Making 4 x 3 plots, each showing relationships with

```{r compplots}

dat.long <- dat %>% 
  pivot_longer(cols = 1:12, names_to = "ES_response", values_to = "ES_value") %>%
  pivot_longer(cols = c("ntr_pland", "ann_agr_pland"), names_to = "Landuse",
               values_to = "Landuse_value") %>%
  select(ES_response, ES_value, Landuse, Landuse_value) %>%
  mutate(ES = gsub("delta_|_base|_max.margin","", ES_response)) %>%
  mutate(ES = gsub("sed$","sed_export", ES)) %>%
  mutate(response_type =
           gsub("sed|export|n_|pollinators|qb|_|max","",ES_response))

# fit some gams
dat.gam <- dat.long %>% filter(Landuse == "ntr_pland") %>% 
  group_by(response_type, ES) %>%
  nest() %>%
  # Fit GAM
  mutate(Mod = map(data, ~mgcv::gam(ES_value ~ s(Landuse_value), data = .x))) %>%
  # Get R2
  mutate(R2 = map_dbl(Mod, ~round(summary(.x)$r.sq, 2)))

ggplot(dat.long %>% filter(Landuse == "ntr_pland"),
       aes(Landuse_value, ES_value)) + 
  geom_point() + geom_smooth(method="gam") +
  ggh4x::facet_grid2(response_type ~ ES, 
                                    scales = "free", independent = "y") + 
  egg::theme_article() + 
  # Add label
  geom_label(data = dat.gam, 
             aes(x = Inf, y = Inf, 
                 label = paste("R2 = ", R2, sep = " ")),
             hjust = 1, vjust = 1)
```

# Model performance
*Pollinators:*

   - Base

     r2 = `r round(pollinators_base.rf$r.squared, 3)`
     OOB MSE = `r round(pollinators_base.rf$prediction.error, 3)`

  - Margin

     r2 = `r round(pollinators_margin.rf$r.squared, 3)`
     OOB MSE = `r round(pollinators_margin.rf$prediction.error, 3)`

*Nitrogen export:*

   - Base

     r2 = `r round(n_export_base.rf$r.squared, 3)`
     OOB MSE = `r round(n_export_base.rf$prediction.error, 3)`

  - Margin

     r2 = `r round(n_export_margin.rf$r.squared, 3)`
     OOB MSE = `r round(n_export_margin.rf$prediction.error, 3)`

*Sediment export:*

   - Base

     r2 = `r round(sed_export_base.rf$r.squared, 3)`
     OOB MSE = `r round(sed_export_base.rf$prediction.error, 3)`

  - Margin

     r2 = `r round(sed_export_margin.rf$r.squared, 3)`
     OOB MSE = `r round(sed_export_margin.rf$prediction.error, 3)`

*Baseflow:*

   - Base

     r2 = `r round(qb_base.rf$r.squared, 3)`
     OOB MSE = `r round(qb_base.rf$prediction.error, 3)`

  - Margin

     r2 = `r round(qb_margin.rf$r.squared, 3)`
     OOB MSE = `r round(qb_margin.rf$prediction.error, 3)`

# Variable importance

```{r importance, fig.height=16, fig.width=9}
library(ggpattern)

metric.descr <- read.csv("data/lsm_description.csv")

imp.df <- bind_rows(
  pollinators_base.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Pollinators", model = "base"),
  pollinators_margin.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Pollinators", model = "margin"),
  n_export_base.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "N export", model = "base"),
  n_export_margin.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "N export", model = "margin"),
  sed_export_base.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Sed export", model = "base"),
  sed_export_margin.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Sed export", model = "margin"),
  qb_base.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Baseflow", model = "base"),
  qb_margin.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Baseflow", model = "margin")
) %>%
  mutate(sig = ifelse(pvalue <0.05, TRUE, FALSE)) %>%
  mutate(metric2 = tidytext::reorder_within(metric, importance,
                                           within = list(ES, model))) %>%
  left_join(metric.descr %>% select(lsm, type), by = c("metric"="lsm")) %>%
  mutate(type = case_match(
    metric,
    "field_pland" ~ "area and edge metric",
    "field_frac_mn" ~ "shape metric",
    "field_frac_cv" ~ "shape metric",
    "prn_ann_ratio" ~ "area and edge metric",
    .default = type
  ))

ggplot(imp.df, aes(metric2, importance, fill = type, pattern = sig)) + 
  geom_bar_pattern(stat="identity",
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,
                   pattern_spacing = 0.025,
                   pattern_key_scale_factor = 0.6) + 
  tidytext::scale_x_reordered() +
  coord_flip() +
  ggh4x::facet_grid2(ES~model, scales = "free", independent = "all") +
  scale_pattern_manual(values = c("TRUE" = "none", "FALSE" = "stripe")) +
  egg::theme_article() +
  theme(legend.position = "top") + 
  guides(pattern = guide_legend(override.aes = list(fill = "white")),
         fill = guide_legend(override.aes = list(pattern = "none")))

```

*Significant metrics count*

```{r}

sigct <- imp.df %>% filter(sig==TRUE)

ggplot(sigct, aes(fct_infreq(metric), fill = type)) +
  scale_x_discrete(limits=rev) +
  geom_bar() + coord_flip() + labs(x="metric") + egg::theme_article()

```

# Interactions
```{r int_load}
library(iml)
library(gridExtra)
load("data/ES-RF2-preds.RDA")
```

## Pollinator base interactions

```{r polbase_int}
grid.arrange(
  int.polbase %>% plot() + ggtitle("Interaction strength") + egg::theme_article() + theme(plot.title.position = "plot", axis.title.y = element_blank()),
  int.polbase.ntr_pland %>% plot() + ggtitle("Pairwise interactions with strongest") + egg::theme_article() + theme(plot.title.position = "plot", axis.title.y = element_blank()),
  ncol = 2
)
```

## Pollinators margins interactions

```{r polmarg_int}
grid.arrange(
  int.polmarg %>% plot() + ggtitle("Interaction strength") + egg::theme_article() + theme(plot.title.position = "plot", axis.title.y = element_blank()),
  int.polmarg.ntr_pland %>% plot() + ggtitle("Pairwise interactions with strongest") + egg::theme_article() + theme(plot.title.position = "plot", axis.title.y = element_blank()),
  ncol = 2
)
```

## N export base interactions

```{r nexpbase_int}
grid.arrange(
  int.nexpbase %>% plot() + ggtitle("Interaction strength") + egg::theme_article() + theme(plot.title.position = "plot", axis.title.y = element_blank()),
  int.nexpbase.ntr_pland %>% plot() + ggtitle("Pairwise interactions with strongest") + egg::theme_article() + theme(plot.title.position = "plot", axis.title.y = element_blank()),
  ncol = 2
)
```

## N export margin interactions

```{r nexpmarg_int}
grid.arrange(
  int.nexpmarg %>% plot() + ggtitle("Interaction strength") + egg::theme_article() + theme(plot.title.position = "plot", axis.title.y = element_blank()),
  int.nexpmarg.ntr_pland %>% plot() + ggtitle("Pairwise interactions with strongest") + egg::theme_article() + theme(plot.title.position = "plot", axis.title.y = element_blank()),
  ncol = 2
)
```

## Sediment export base interactions

```{r sedbase_int}
grid.arrange(
  int.sedexpbase %>% plot() + ggtitle("Interaction strength") + egg::theme_article() + theme(plot.title.position = "plot", axis.title.y = element_blank()),
  int.sedexpbase.ann_agr_ed %>% plot() + ggtitle("Pairwise interactions with strongest") + egg::theme_article() + theme(plot.title.position = "plot", axis.title.y = element_blank()),
  ncol = 2
)
```

## Sediment export margin interactions

```{r sedmarg_int}
grid.arrange(
  int.sedexpmarg %>% plot() + ggtitle("Interaction strength") + egg::theme_article() + theme(plot.title.position = "plot", axis.title.y = element_blank()),
  int.sedexpmarg.ann_agr_ed %>% plot() + ggtitle("Pairwise interactions with strongest") + egg::theme_article() + theme(plot.title.position = "plot", axis.title.y = element_blank()),
  ncol = 2
)
```

## Baseflow base interactions

```{r qbbase_int}
grid.arrange(
  int.qbbase %>% plot() + ggtitle("Interaction strength") + egg::theme_article() + theme(plot.title.position = "plot", axis.title.y = element_blank()),
  int.qbbase.field_frac_mn %>% plot() + ggtitle("Pairwise interactions with strongest") + egg::theme_article() + theme(plot.title.position = "plot", axis.title.y = element_blank()),
  ncol = 2
)
```

## Baseflow margin interactions

```{r qbmarg_int}
grid.arrange(
  int.qbmarg %>% plot() + ggtitle("Interaction strength") + egg::theme_article() + theme(plot.title.position = "plot", axis.title.y = element_blank()),
  int.qbmarg.prn_agr_para_mn %>% plot() + ggtitle("Pairwise interactions with strongest") + egg::theme_article() + theme(plot.title.position = "plot", axis.title.y = element_blank()),
  ncol = 2
)
```