### 2025-04-01 Paper_analysis.Rmd
update again

---
title: "Paper analysis"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r load}
library(tidyverse)
library(iml)
library(ranger)
load("data/ES-RF3.RDA")

```

## Patterns with composition

Making 4 x 3 plots, each showing relationships with

```{r compplots}

dat.long <- dat %>% 
  pivot_longer(cols = 1:12, names_to = "ES_response", values_to = "ES_value") %>%
  pivot_longer(cols = c("ntr_pland", "ann_agr_pland"), names_to = "Landuse",
               values_to = "Landuse_value") %>%
  select(ES_response, ES_value, Landuse, Landuse_value) %>%
  mutate(ES = gsub("delta_|_base|_max.margin","", ES_response)) %>%
  mutate(ES = gsub("sed$","sed_export", ES)) %>%
  mutate(response_type =
           gsub("sed|export|n_|pollinators|qb|_|max","",ES_response))

# fit some gams
dat.gam <- dat.long %>% filter(Landuse == "ntr_pland") %>% 
  group_by(response_type, ES) %>%
  nest() %>%
  # Fit GAM
  mutate(Mod = map(data, ~mgcv::gam(ES_value ~ s(Landuse_value), data = .x))) %>%
  # Get R2
  mutate(R2 = map_dbl(Mod, ~round(summary(.x)$r.sq, 2)))

ggplot(dat.long %>% filter(Landuse == "ntr_pland"),
       aes(Landuse_value, ES_value)) + 
  geom_point() + geom_smooth(method="gam") +
  ggh4x::facet_grid2(response_type ~ ES, 
                                    scales = "free", independent = "y") + 
  egg::theme_article() + 
  # Add label
  geom_label(data = dat.gam, 
             aes(x = Inf, y = Inf, 
                 label = paste("R2 = ", R2, sep = " ")),
             hjust = 1, vjust = 1)
```

# Model performance
*Pollinators:*

   - Base

     r2 = `r round(pollinators_base.rf$r.squared, 3)`
     OOB MSE = `r round(pollinators_base.rf$prediction.error, 3)`

  - Margin

     r2 = `r round(pollinators_margin.rf$r.squared, 3)`
     OOB MSE = `r round(pollinators_margin.rf$prediction.error, 3)`

*Nitrogen export:*

   - Base

     r2 = `r round(n_export_base.rf$r.squared, 3)`
     OOB MSE = `r round(n_export_base.rf$prediction.error, 3)`

  - Margin

     r2 = `r round(n_export_margin.rf$r.squared, 3)`
     OOB MSE = `r round(n_export_margin.rf$prediction.error, 3)`

*Sediment export:*

   - Base

     r2 = `r round(sed_export_base.rf$r.squared, 3)`
     OOB MSE = `r round(sed_export_base.rf$prediction.error, 3)`

  - Margin

     r2 = `r round(sed_export_margin.rf$r.squared, 3)`
     OOB MSE = `r round(sed_export_margin.rf$prediction.error, 3)`

*Baseflow:*

   - Base

     r2 = `r round(qb_base.rf$r.squared, 3)`
     OOB MSE = `r round(qb_base.rf$prediction.error, 3)`

  - Margin

     r2 = `r round(qb_margin.rf$r.squared, 3)`
     OOB MSE = `r round(qb_margin.rf$prediction.error, 3)`

# Variable importance

```{r importance, fig.height=16, fig.width=9}
library(ggpattern)

metric.descr <- read.csv("data/lsm_description.csv")

imp.df <- bind_rows(
  pollinators_base.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Pollinators", model = "base"),
  pollinators_margin.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Pollinators", model = "margin"),
  n_export_base.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "N export", model = "base"),
  n_export_margin.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "N export", model = "margin"),
  sed_export_base.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Sed export", model = "base"),
  sed_export_margin.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Sed export", model = "margin"),
  qb_base.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Baseflow", model = "base"),
  qb_margin.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Baseflow", model = "margin")
) %>%
  mutate(sig = ifelse(pvalue <0.05, TRUE, FALSE)) %>%
  mutate(metric2 = tidytext::reorder_within(metric, importance,
                                           within = list(ES, model))) %>%
  left_join(metric.descr %>% select(lsm, type), by = c("metric"="lsm")) %>%
  mutate(type = case_match(
    metric,
    "field_pland" ~ "area and edge metric",
    "field_frac_mn" ~ "shape metric",
    "field_frac_cv" ~ "shape metric",
    "prn_pagr" ~ "area and edge metric",
    .default = type
  ))

# ggplot(imp.df, aes(metric2, importance, fill = type, pattern = sig)) + 
#   geom_bar_pattern(stat="identity",
#                    color = "black", 
#                    pattern_fill = "black",
#                    pattern_angle = 45,
#                    pattern_density = 0.1,
#                    pattern_spacing = 0.025,
#                    pattern_key_scale_factor = 0.6) + 
#   tidytext::scale_x_reordered() +
#   coord_flip() +
#   ggh4x::facet_grid2(ES~model, scales = "free", independent = "all") +
#   scale_pattern_manual(values = c("TRUE" = "none", "FALSE" = "stripe")) +
#   egg::theme_article() +
#   theme(legend.position = "top") + 
#   guides(pattern = guide_legend(override.aes = list(fill = "white")),
#          fill = guide_legend(override.aes = list(pattern = "none")))

```

```{r importance_shap}
load("data/ES-RF3-preds.RDA")   # prediction objects from iml package
load("data/ES-RF3-shapimportance.RDA")
library(shapviz)
library(gridExtra)

```

### Pollinators

#### Permutation based importance

```{r pol_permimport, fig.width=9}

ggplot(imp.df %>% filter(ES=="Pollinators"), aes(metric2, importance, fill = type, pattern = sig)) + 
  geom_bar_pattern(stat="identity",
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,
                   pattern_spacing = 0.025,
                   pattern_key_scale_factor = 0.6) + 
  tidytext::scale_x_reordered() +
  coord_flip() +
  facet_wrap(~model, scales = "free") +
  scale_pattern_manual(values = c("TRUE" = "none", "FALSE" = "stripe")) +
  egg::theme_article() +
  theme(legend.position = "top") + 
  guides(pattern = guide_legend(override.aes = list(fill = "white")),
         fill = guide_legend(override.aes = list(pattern = "none")))

```

#### Shapley importance

```{r shap_pol, fig.width=9}
cowplot::plot_grid(
  sv_importance(ex.polbase.glb) + egg::theme_article(),
  sv_importance(ex.polmarg.glb) + egg::theme_article(),
  ncol = 2
)

```

#### Base pollinator model Shapley dependence plots

The two major contributors. Colors are the strongest interactor with that variable.

```{r shap_pol_baseshap, fig.width=9}
grid.arrange(
  sv_dependence(ex.polbase.glb, v="ntr_pland") + egg::theme_article(),
  sv_dependence(ex.polbase.glb, v="ann_agr_pland") + egg::theme_article(), 
  ncol = 2
)

```

**_Compare to PDP and ALE_**

```{r pol_base_ale, fig.width=9}

polbase.ntr_pland.pdp = FeatureEffect$new(pred.pollinators_base, 
                                          feature = "ntr_pland",
                                          method = "pdp")
polbase.ann_agr_pland.pdp = FeatureEffect$new(pred.pollinators_base, 
                                          feature = "ann_agr_pland",
                                          method = "pdp")

polbase.ntr_pland.ale = FeatureEffect$new(pred.pollinators_base, 
                                          feature = "ntr_pland",
                                          method = "ale")
polbase.ann_agr_pland.ale = FeatureEffect$new(pred.pollinators_base, 
                                          feature = "ann_agr_pland",
                                          method = "ale")

cowplot::plot_grid(
  polbase.ntr_pland.pdp$plot() + ggtitle("ntr_pland PDP") + egg::theme_article(),
  polbase.ntr_pland.ale$plot() + ggtitle("ntr_pland ALE") + egg::theme_article(),

  polbase.ann_agr_pland.pdp$plot() + ggtitle("ann_agr_pland PDP") + egg::theme_article(),
  polbase.ann_agr_pland.pdp$plot() + ggtitle("ann_agr_pland ALE") + egg::theme_article(),
  ncol = 2, align = "v"
)

```

#### Margin pollinator model Shapley dependence plot

Colors are the strongest interactor with that variable.

```{r shap_pol_margshap, fig.width=9}
cowplot::plot_grid(
  sv_dependence(ex.polmarg.glb, v="ntr_pland") + egg::theme_article(),
  sv_dependence(ex.polmarg.glb, v="ann_agr_pland") + egg::theme_article(),
  ncol = 2, align = "v"
)

```

**_Compare to PDP and ALE_**

```{r pol_marg_ale, fig.width=9}

polmarg.ntr_pland.pdp = FeatureEffect$new(pred.pollinators_margin, 
                                          feature = "ntr_pland",
                                          method = "pdp")
polmarg.ann_agr_pland.pdp = FeatureEffect$new(pred.pollinators_margin, 
                                          feature = "ann_agr_pland",
                                          method = "pdp")

polmarg.ntr_pland.ale = FeatureEffect$new(pred.pollinators_margin, 
                                          feature = "ntr_pland",
                                          method = "ale")
polmarg.ann_agr_pland.ale = FeatureEffect$new(pred.pollinators_margin, 
                                          feature = "ann_agr_pland",
                                          method = "ale")

cowplot::plot_grid(
  polmarg.ntr_pland.pdp$plot() + ggtitle("ntr_pland PDP") + egg::theme_article(),
  polmarg.ntr_pland.ale$plot() + ggtitle("ntr_pland ALE") + egg::theme_article(),

  polmarg.ann_agr_pland.pdp$plot() + ggtitle("ann_agr_pland PDP") + egg::theme_article(),
  polmarg.ann_agr_pland.ale$plot() + ggtitle("ann_agr_pland ALE") + egg::theme_article(),
  ncol = 2, align = "v"
)

```

### N Export

### Permutation based importance

```{r nexp_permimport, fig.width=9}

ggplot(imp.df %>% filter(ES=="N export"), aes(metric2, importance, fill = type, pattern = sig)) + 
  geom_bar_pattern(stat="identity",
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,
                   pattern_spacing = 0.025,
                   pattern_key_scale_factor = 0.6) + 
  tidytext::scale_x_reordered() +
  coord_flip() +
  facet_wrap(~model, scales = "free") +
  scale_pattern_manual(values = c("TRUE" = "none", "FALSE" = "stripe")) +
  egg::theme_article() +
  theme(legend.position = "top") + 
  guides(pattern = guide_legend(override.aes = list(fill = "white")),
         fill = guide_legend(override.aes = list(pattern = "none")))

```

#### Shapley importance

```{r shap_nexp, fig.width=9}

cowplot::plot_grid(
  sv_importance(ex.nexpbase.glb) + egg::theme_article(), 
  sv_importance(ex.nexpmarg.glb) + egg::theme_article(),
  ncol = 2
)

```

#### Base N export shapley dependence plots

Three major -ish contributors.

```{r shap_nexp_baseshap, fig.width=9}

cowplot::plot_grid(
  sv_dependence(ex.nexpbase.glb, v="ntr_pland") + egg::theme_article(),
  sv_dependence(ex.nexpbase.glb, v="ann_agr_pland") + egg::theme_article(), 
  sv_dependence(ex.nexpbase.glb, v="ntr_ed") + egg::theme_article(), 
  ncol = 2, align = "v"
)

```

**_Compare to PDP and ALE_**

```{r nexp_base_ale, fig.width=9}

nexpbase.ntr_pland.pdp = FeatureEffect$new(pred.n_export_base, 
                                          feature = "ntr_pland",
                                          method = "pdp")
nexpbase.ann_agr_pland.pdp = FeatureEffect$new(pred.n_export_base, 
                                          feature = "ann_agr_pland",
                                          method = "pdp")
nexpbase.ntr_ed.pdp = FeatureEffect$new(pred.n_export_base, 
                                          feature = "ntr_ed",
                                          method = "pdp")

nexpbase.ntr_pland.ale = FeatureEffect$new(pred.n_export_base, 
                                          feature = "ntr_pland",
                                          method = "ale")
nexpbase.ann_agr_pland.ale = FeatureEffect$new(pred.n_export_base, 
                                          feature = "ann_agr_pland",
                                          method = "ale")
nexpbase.ntr_ed.ale = FeatureEffect$new(pred.n_export_base, 
                                          feature = "ntr_ed",
                                          method = "ale")

cowplot::plot_grid(
  nexpbase.ntr_pland.pdp$plot() + ggtitle("ntr_pland PDP") + egg::theme_article(),
  nexpbase.ntr_pland.ale$plot() + ggtitle("ntr_pland ALE") + egg::theme_article(),

  nexpbase.ann_agr_pland.pdp$plot() + ggtitle("ann_agr_pland PDP") + egg::theme_article(),
  nexpbase.ann_agr_pland.ale$plot() + ggtitle("ann_agr_pland ALE") + egg::theme_article(),

  nexpbase.ntr_ed.pdp$plot() + ggtitle("ntr_ed PDP") + egg::theme_article(),
  nexpbase.ntr_ed.ale$plot() + ggtitle("ntr_ed ALE") + egg::theme_article(),
  ncol = 2, align = "v"
)

```

#### Margin N export Shapley dependence plots

Three major contributors.

```{r shap_nexp_margshap, fig.width=9}
cowplot::plot_grid(
  sv_dependence(ex.nexpmarg.glb, v="ntr_pland") + egg::theme_article(),
  sv_dependence(ex.nexpmarg.glb, v="ann_agr_pland") + egg::theme_article(), 
  sv_dependence(ex.nexpmarg.glb, v="ntr_ed") + egg::theme_article(), 
  ncol = 2, align = "v"
)

```

**_Compare to PDP and ALE_**

```{r nexp_marg_ale, fig.width=9}

nexpmarg.ntr_pland.pdp = FeatureEffect$new(pred.n_export_margin, 
                                          feature = "ntr_pland",
                                          method = "pdp")
nexpmarg.ann_agr_pland.pdp = FeatureEffect$new(pred.n_export_margin, 
                                          feature = "ann_agr_pland",
                                          method = "pdp")
nexpmarg.ntr_ed.pdp = FeatureEffect$new(pred.n_export_margin, 
                                          feature = "ntr_ed",
                                          method = "pdp")

nexpmarg.ntr_pland.ale = FeatureEffect$new(pred.n_export_margin, 
                                          feature = "ntr_pland",
                                          method = "ale")
nexpmarg.ann_agr_pland.ale = FeatureEffect$new(pred.n_export_margin, 
                                          feature = "ann_agr_pland",
                                          method = "ale")
nexpmarg.ntr_ed.ale = FeatureEffect$new(pred.n_export_margin, 
                                          feature = "ntr_ed",
                                          method = "ale")

cowplot::plot_grid(
  nexpmarg.ntr_pland.pdp$plot() + ggtitle("ntr_pland PDP") + egg::theme_article(),
  nexpmarg.ntr_pland.ale$plot() + ggtitle("ntr_pland ALE") + egg::theme_article(),

  nexpmarg.ann_agr_pland.pdp$plot() + ggtitle("ann_agr_pland PDP") + egg::theme_article(),
  nexpmarg.ann_agr_pland.ale$plot() + ggtitle("ann_agr_pland ALE") + egg::theme_article(),

  nexpmarg.ntr_ed.pdp$plot() + ggtitle("ntr_ed PDP") + egg::theme_article(),
  nexpmarg.ntr_ed.ale$plot() + ggtitle("ntr_ed ALE") + egg::theme_article(),
  ncol = 2, align = "v"
)

```

### Sediment Export

### Permutation based importance

```{r sed_permimport, fig.width=9}

ggplot(imp.df %>% filter(ES=="Sed export"), aes(metric2, importance, fill = type, pattern = sig)) + 
  geom_bar_pattern(stat="identity",
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,
                   pattern_spacing = 0.025,
                   pattern_key_scale_factor = 0.6) + 
  tidytext::scale_x_reordered() +
  coord_flip() +
  facet_wrap(~model, scales = "free") +
  scale_pattern_manual(values = c("TRUE" = "none", "FALSE" = "stripe")) +
  egg::theme_article() +
  theme(legend.position = "top") + 
  guides(pattern = guide_legend(override.aes = list(fill = "white")),
         fill = guide_legend(override.aes = list(pattern = "none")))

```

#### Shapley importance

```{r shap_sed, fig.width=9}
cowplot::plot_grid(
  sv_importance(ex.sedbase.glb) + egg::theme_article(),
  sv_importance(ex.sedmarg.glb) + egg::theme_article(),
  ncol = 2
)

```

#### Base sediment export Shapley dependence plots

One major contributor

```{r shap_sed_baseshap, fig.width=9}
sv_dependence(ex.sedbase.glb, v="ann_agr_ed") + egg::theme_article()

```

**_Compare to PDP and ALE_**

```{r sedexp_base_ale, fig.width=9}

sedbase.ann_agr_ed.pdp = FeatureEffect$new(pred.sed_export_base, 
                                          feature = "ann_agr_ed",
                                          method = "pdp")

sedbase.ann_agr_ed.ale = FeatureEffect$new(pred.sed_export_base, 
                                          feature = "ann_agr_ed",
                                          method = "ale")

cowplot::plot_grid(
  sedbase.ann_agr_ed.pdp$plot() + ggtitle("ann_agr_ed PDP") +
    egg::theme_article(),
  sedbase.ann_agr_ed.ale$plot() + ggtitle("ann_agr_ed ALE") +
    egg::theme_article(),

  ncol = 2, align = "v"
)

```

#### Margin sediment runoff Shapley dependence plots

One or two major contributors. `field_frac_mn` is the strongest interactor with `ann_agr_ed`, and vice versa, while `field_frac_mn` does not stand out as strongly in the permutation results.

```{r shap_sed_margshap, fig.width=9}
cowplot::plot_grid(
  sv_dependence(ex.sedmarg.glb, v="ann_agr_ed") + egg::theme_article(),
  sv_dependence(ex.sedmarg.glb, v="field_frac_mn") + egg::theme_article(),
  ncol=2, align = "v"
)

```

**_Compare to PDP and ALE_**

```{r sedexp_marg_ale, fig.width=9}

sedmarg.ann_agr_ed.pdp = FeatureEffect$new(pred.sed_export_margin, 
                                          feature = "ann_agr_ed",
                                          method = "pdp")
sedmarg.field_frac_mn.pdp = FeatureEffect$new(pred.sed_export_margin, 
                                          feature = "field_frac_mn",
                                          method = "pdp")

sedmarg.ann_agr_ed.ale = FeatureEffect$new(pred.sed_export_margin, 
                                          feature = "ann_agr_ed",
                                          method = "ale")
sedmarg.field_frac_mn.ale = FeatureEffect$new(pred.sed_export_margin, 
                                          feature = "field_frac_mn",
                                          method = "ale")

cowplot::plot_grid(
  sedmarg.ann_agr_ed.pdp$plot() + ggtitle("ann_agr_ed PDP") +
    egg::theme_article(),
  sedmarg.ann_agr_ed.ale$plot() + ggtitle("ann_agr_ed ALE") +
    egg::theme_article(),

  sedmarg.field_frac_mn.pdp$plot() + ggtitle("field_frac_mn PDP") +
    egg::theme_article(),
  sedmarg.field_frac_mn.ale$plot() + ggtitle("field_frac_mn ALE") +
    egg::theme_article(),

  ncol = 2, align = "v"
)

```

### Water recharge

### Permutation based importance

```{r qb_permimport, fig.width=9}

ggplot(imp.df %>% filter(ES=="Baseflow"), aes(metric2, importance, fill = type, pattern = sig)) + 
  geom_bar_pattern(stat="identity",
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,
                   pattern_spacing = 0.025,
                   pattern_key_scale_factor = 0.6) + 
  tidytext::scale_x_reordered() +
  coord_flip() +
  facet_wrap(~model, scales = "free") +
  scale_pattern_manual(values = c("TRUE" = "none", "FALSE" = "stripe")) +
  egg::theme_article() +
  theme(legend.position = "top") + 
  guides(pattern = guide_legend(override.aes = list(fill = "white")),
         fill = guide_legend(override.aes = list(pattern = "none")))

```

#### Shapley importance

```{r shap_qb, fig.width=9}
cowplot::plot_grid(
  sv_importance(ex.qbbase.glb) + egg::theme_article(),
  sv_importance(ex.qbmarg.glb) + egg::theme_article(),
  ncol = 2
)

```

#### Base water runoff Shapley depednence plot

One major contributor.

```{r shap_qb_baseshap, fig.width=9}
cowplot::plot_grid(
  sv_dependence(ex.qbbase.glb, v="ann_agr_cohesion") + egg::theme_article(),
  sv_dependence(ex.qbbase.glb, v="field_frac_mn") + egg::theme_article(),
  ncol = 2, align = "v"
)

```

**_Compare to PDP and ALE_**

```{r qb_base_ale, fig.width=9}

qbbase.ann_agr_cohesion.pdp = FeatureEffect$new(pred.qb_base, 
                                             feature = "ann_agr_cohesion",
                                             method = "pdp")

sedbase.ann_agr_cohesion.ale = FeatureEffect$new(pred.qb_base, 
                                              feature = "ann_agr_cohesion",
                                              method = "ale")

qbbase.field_frac_mn.pdp = FeatureEffect$new(pred.qb_base, 
                                             feature = "field_frac_mn",
                                             method = "pdp")

sedbase.field_frac_mn.ale = FeatureEffect$new(pred.qb_base, 
                                              feature = "field_frac_mn",
                                              method = "ale")

cowplot::plot_grid(
  qbbase.ann_agr_cohesion.pdp$plot() + ggtitle("ann_agr_cohesion PDP") +
    egg::theme_article(),
  sedbase.ann_agr_cohesion.ale$plot() + ggtitle("ann_agr_cohesion ALE") +
    egg::theme_article(),

  qbbase.field_frac_mn.pdp$plot() + ggtitle("field_frac_mn PDP") +
    egg::theme_article(),
  sedbase.field_frac_mn.ale$plot() + ggtitle("field_frac_mn ALE") +
    egg::theme_article(),

  ncol = 2, align = "v"
)

```

#### Margin water runoff Shapley dependence plots

One or two major contributors. Differs a little from the permutation results from above. In the permutation results, `prn_agr_para_mn` does not have as high importance so that it isn't separated from other variables.

```{r shap_qb_margshap, fig.width=9}
cowplot::plot_grid(
  sv_dependence(ex.qbmarg.glb, v="ann_agr_cohesion") + egg::theme_article(),
  sv_dependence(ex.qbmarg.glb, v="prn_pagr") + egg::theme_article(),
  sv_dependence(ex.qbmarg.glb, v="prn_agr_para_mn") + egg::theme_article(),
  ncol=2, align = "v"
)

```

**_Compare to PDP and ALE_**

```{r qb_marg_ale, fig.width=9}

qbmarg.ann_agr_cohesion.pdp = FeatureEffect$new(pred.qb_margin, 
                                             feature = "ann_agr_cohesion",
                                             method = "pdp")
qbmarg.prn_pagr.pdp = FeatureEffect$new(pred.qb_margin, 
                                             feature = "prn_pagr",
                                             method = "pdp")
qbmarg.prn_agr_para_mn.pdp = FeatureEffect$new(pred.qb_margin, 
                                             feature = "prn_agr_para_mn",
                                             method = "pdp")

qbmarg.ann_agr_cohesion.ale = FeatureEffect$new(pred.qb_margin, 
                                              feature = "ann_agr_cohesion",
                                              method = "ale")
qbmarg.prn_pagr.ale = FeatureEffect$new(pred.qb_margin, 
                                              feature = "prn_pagr",
                                              method = "ale")
qbmarg.prn_agr_para_mn.ale = FeatureEffect$new(pred.qb_margin, 
                                              feature = "prn_agr_para_mn",
                                              method = "ale")

cowplot::plot_grid(
  qbmarg.ann_agr_cohesion.pdp$plot() + ggtitle("ann_agr_cohesion PDP") + ylab("PDP") +
    egg::theme_article(),
  qbmarg.ann_agr_cohesion.ale$plot() + ggtitle("ann_agr_cohesion ALE") + ylab("ALE") +
    egg::theme_article(),

  qbmarg.prn_pagr.pdp$plot() + ggtitle("prn_pagr PDP") + ylab("PDP") +
    egg::theme_article(),
  qbmarg.prn_pagr.ale$plot() + ggtitle("prn_pagr ALE") + ylab("ALE") +
    egg::theme_article(),

  qbmarg.prn_agr_para_mn.pdp$plot() + ggtitle("prn_agr_para_mn PDP") + ylab("PDP") +
    egg::theme_article(),
  qbmarg.prn_agr_para_mn.ale$plot() + ggtitle("prn_agr_para_mn ALE") + ylab("ALE") +
    egg::theme_article(),

  ncol = 2, align = "v"
)

```

#### Correlation of important variables

```{r paircor}
selectedvar <- c("ntr_pland", "ann_agr_pland", "ntr_ed", 
                 "field_frac_mn", "prn_pagr", "prn_agr_para_mn",
                 "ann_agr_ed", "ann_agr_cohesion"
                 )

# selectedvar.base <- c("ntr_pland", "ann_agr_pland",  "ntr_ed", 
#                       "ann_agr_ed", "field_frac_mn")
# 
# selectedvar.margin <- c("ntr_pland", "ann_agr_pland",  "ntr_ed",
#                         "ann_agr_ed", "field_frac_mn",
#                         "prn_ann_ratio", "prn_agr_para_mn")

dat.selectedvar <- dat %>% select(selectedvar)
psych::pairs.panels(dat.selectedvar)

```

### SHAP force plots

I am plotting the combined SHAP force values for each watershed together. The important features identified above are colored while the rest are left grey. To plot watersheds on a common axis and visualize tradeoffs, I used PCA to identify the two major axes. The first PC axis for both baseline and margin results show a tradeoff between N export (and improvement in N export) and pollinator abundance (pollinator improvement). The second PC shows sediment export and baseflow changing together while with margin improvement are in the same direction.

```{r force_fcns}
source("R/utils-shap.r")
focus.vars <- function(x.features){
  dplyr::case_match(
    x.features,
    "ntr_pland" ~ "Natural land cover",
    "ann_agr_pland" ~ "Annual ag. land cover",
    "field_frac_mn" ~ "Field fractal dimension",
    "prn_pagr" ~ "Agriculture % perennial",
    "prn_agr_para_mn" ~ "Perennial ag. perimeter:area",
    "ann_agr_ed" ~ "Annual ag. edge density",
    "ann_agr_cohesion" ~ "Annual ag. cohesion",
    .default = "Other features"
  )
}

feature.colors <- c(
  "Natural land cover" = "#33a02c",
  "Annual ag. land cover" = "#fdbf6f",
  "Field fractal dimension" = "#56B4E9",
  "Agriculture % perennial" = "#CC79A7",
  "Perennial ag. perimeter:area" = "#D55E00",
  "Annual ag. edge density" = "#F0E442",
  "Annual ag. cohesion" = "#0072B2",
  "Other features" = "gray80"
)

```

```{r make_pca_baseoutcomes_forces, include = FALSE}

base_esoutcome <- dat %>% select(pollinators_base:sed_export_base) %>%
  mutate(sed_export_base = sed_export_base,
         n_export_base = n_export_base)

base_pca <- prcomp(base_esoutcome,
                   scale=TRUE, center=TRUE)

ntr_class <- cut(dat$ntr_pland, breaks = c(0, 10, 20, 30 , 50, 100),
                 labels = c("5-10%", "10-20%", "20-30%", "30-50%", "50-90%"))

biplot_base_pca <- ggbiplot::ggbiplot(base_pca, groups = ntr_class) + 
  expand_limits(x=c(-2.5,2.5)) +
  scale_color_manual(name="Natural land cover",
                     values = c("#d7191c", "#fdae61", "#ffffbf", "#abdda4", "#2b83ba")) +
  egg::theme_article() + 
  theme(legend.position = "inside", legend.position.inside = c(0.2,0.12))

biplot_base_arrows <- biplot_base_pca$layers[[3]]$data
biplot_base_arrows$varname <- c("Pollinators", "Baseflow", "N export", "Sediment export")

# biplot_base_pca$layers[[1]]$aes_params$colour = "gray"
biplot_base_pca$layers[[1]]$aes_params$size = 1
biplot_base_pca$layers[[3]] <- geom_label(aes(x = xvar, y = yvar, label = varname,
                                  angle = angle, hjust = hjust), 
                              label.size = NA, size = 3,
                              data = biplot_base_arrows, 
                              fill = '#dddddd80')
biplot_base_pca$layers <- cbind(biplot_base_pca$layers[[3]], biplot_base_pca$layers)

psych::pairs.panels(dat%>% select(pollinators_base:sed_export_base))

polbase_pc1 <- makeforcedata.orderby(ex.polbase$shapley_values,
                                     predict(base_pca)[,"PC1"],
                               rename.fn = focus.vars, other.var = "Other features")

nexpbase_pc1 <- makeforcedata.orderby(ex.nexpbase$shapley_values,
                                  predict(base_pca)[,"PC1"],
                               rename.fn = focus.vars, other.var = "Other features")

sedbase_pc2 <- makeforcedata.orderby(ex.sedbase$shapley_values,
                                     predict(base_pca)[,"PC2"],
                               rename.fn = focus.vars, other.var = "Other features")

qbbase_pc2 <- makeforcedata.orderby(ex.qbbase$shapley_values,
                                    predict(base_pca)[,"PC2"],
                               rename.fn = focus.vars, other.var = "Other features")

polbase_pc1_force <- ggplot(polbase_pc1, aes(rowid, value, fill = feature2)) +
  geom_col(width = 1) + egg::theme_article() +
  scale_fill_manual(values = feature.colors, guide = guide_legend(reverse=TRUE)) +
  xlab("PC1") + ylab("SHAP values") + ggtitle("Baseline pollinator abundance") +
  theme(axis.ticks.x = element_blank(), axis.text.x = element_blank(),
        legend.position = "none")

nexpbase_pc1_force <- ggplot(nexpbase_pc1, aes(rowid, value, fill = feature2)) +
  geom_col(width = 1) + egg::theme_article() +
  scale_fill_manual(values = feature.colors, guide = guide_legend(reverse=TRUE)) +
  xlab("PC1") + ylab("SHAP values") + ggtitle("Baseline N export") +
  theme(axis.ticks.x = element_blank(), axis.text.x = element_blank(),
        legend.position = "none")

sedbase_pc2_force <- ggplot(sedbase_pc2, aes(rowid, value, fill = feature2)) +
  geom_col(width = 1) + egg::theme_article() +
  scale_fill_manual(values = feature.colors, guide = guide_legend(reverse=TRUE)) +
  xlab("PC2") + ylab("SHAP values") + ggtitle("Baseline sediment export") +
  theme(axis.ticks.x = element_blank(), axis.text.x = element_blank(),
        legend.position = "none")

qbbase_pc2_force <- ggplot(qbbase_pc2, aes(rowid, value, fill = feature2)) +
  geom_col(width = 1) + egg::theme_article() +
  scale_fill_manual(values = feature.colors, guide = guide_legend(reverse=TRUE)) +
  xlab("PC2") + ylab("SHAP values") + ggtitle("Baseline baseflow") +
  theme(axis.ticks.x = element_blank(), axis.text.x = element_blank(),
        legend.position = "none")

base_forcestackPC1 <- cowplot::plot_grid(polbase_pc1_force, nexpbase_pc1_force, align='v',
                                         labels = c("B","C"), ncol=1) %>% 
  ggplotGrob()
base_forcestackPC2 <- cowplot::plot_grid(sedbase_pc2_force, qbbase_pc2_force, align='v',
                                         labels = c("D","E"), ncol=1) %>%
  ggplotGrob()

# make a legend

baselegend_dummy <- ggplotGrob(ggplot(polbase_pc1, aes(rowid, value, fill = feature2)) +
  geom_col(width = 1) + egg::theme_article() +
  scale_fill_manual(values = feature.colors, 
                    guide = guide_legend(reverse=TRUE)) +
  theme(legend.position = "bottom", legend.title = element_blank()))$grobs

base_legend <- baselegend_dummy[[which(sapply(baselegend_dummy, function(x)
      x$name) == "guide-box")]]

base_plots <- cowplot::plot_grid(base_forcestackPC1, base_forcestackPC2, ncol = 2)

base_plots_legend <- grid.arrange(base_plots, base_legend, 
                                  layout_matrix = 
                                    rbind(
                                      c(1,1,1,1,1,1),
                                      c(1,1,1,1,1,1),
                                      c(1,1,1,1,1,1),
                                      c(1,1,1,1,1,1),
                                      c(1,1,1,1,1,1),
                                      c(1,1,1,1,1,1),
                                      c(1,1,1,1,1,1),
                                      c(1,1,1,1,1,1),
                                      c(2,2,2,2,2,2)
                                    ))

base_plots_all <- cowplot::plot_grid(biplot_base_pca, base_plots_legend, 
                                     labels = c("A",""), rel_widths = c(1,2.2))

# ggsave("Figs/base_pcashap.png", base_plots_all, width = 13, height = 5.5, bg = "white", dpi = 300)
```

```{r plot_pca_baseoutcomes, fig.width = 12, fig.height = 5}
base_plots_all
```

```{r make_pca_marginoutcomes, include=FALSE}

marg_esoutcome <- dat %>% select(delta_pollinators_max.margin:delta_sed_max.margin)

marg_pca <- prcomp(marg_esoutcome,
                   scale=TRUE, center=TRUE)

ntr_class <- cut(dat$ntr_pland, breaks = c(0, 10, 20, 30 , 50, 100),
                 labels = c("5-10%", "10-20%", "20-30%", "30-50%", "50-90%"))

biplot_marg_pca <- ggbiplot::ggbiplot(marg_pca, groups = ntr_class) + 
  expand_limits(x=c(-3,3)) +
  scale_color_manual(name="Natural land cover",
                     values = c("#d7191c", "#fdae61", "#ffffbf", "#abdda4", "#2b83ba")) +
  guides(color = guide_legend(label.position = "left")) +
  egg::theme_article() + 
  theme(legend.position = "inside", legend.position.inside = c(0.8,0.85), legend.title.align=1)

biplot_marg_arrows <- biplot_marg_pca$layers[[3]]$data
biplot_marg_arrows$varname <- c("\u0394 Pollinators", "\u0394 Baseflow", "N reduction", "Sediment reduction")

biplot_marg_pca$layers[[1]]$aes_params$size = 1
biplot_marg_pca$layers[[3]] <- geom_label(aes(x = xvar, y = yvar, label = varname,
                                  angle = angle, hjust = hjust), 
                              label.size = NA, size = 3,
                              data = biplot_marg_arrows, 
                              fill = '#dddddd80')
biplot_marg_pca$layers <- cbind(biplot_marg_pca$layers[[3]], biplot_marg_pca$layers)

polmarg_pc1 <- makeforcedata.orderby(ex.polmarg$shapley_values,
                                     predict(marg_pca)[,"PC1"],
                               rename.fn = focus.vars, other.var = "Other features")

nexpmarg_pc1 <- makeforcedata.orderby(ex.nexpmarg$shapley_values,
                                  predict(marg_pca)[,"PC1"],
                               rename.fn = focus.vars, other.var = "Other features")

sedmarg_pc2 <- makeforcedata.orderby(ex.sedmarg$shapley_values,
                                     predict(marg_pca)[,"PC2"],
                               rename.fn = focus.vars, other.var = "Other features")

qbmarg_pc2 <- makeforcedata.orderby(ex.qbmarg$shapley_values,
                                    predict(marg_pca)[,"PC2"],
                               rename.fn = focus.vars, other.var = "Other features")

polmarg_pc1_force <- ggplot(polmarg_pc1, aes(rowid, value, fill = feature2)) +
  geom_col(width = 1) + egg::theme_article() +
  scale_fill_manual(values = feature.colors, guide = guide_legend(reverse=TRUE)) +
  xlab("PC1") + ylab("SHAP values") + ggtitle("Change in pollinator abundance") +
  theme(axis.ticks.x = element_blank(), axis.text.x = element_blank(),
        legend.position = "none")

nexpmarg_pc1_force <- ggplot(nexpmarg_pc1, aes(rowid, value, fill = feature2)) +
  geom_col(width = 1) + egg::theme_article() +
  scale_fill_manual(values = feature.colors, guide = guide_legend(reverse=TRUE)) +
  xlab("PC1") + ylab("SHAP values") + ggtitle("N export reduction") +
  theme(axis.ticks.x = element_blank(), axis.text.x = element_blank(),
        legend.position = "none")

sedmarg_pc2_force <- ggplot(sedmarg_pc2, aes(rowid, value, fill = feature2)) +
  geom_col(width = 1) + egg::theme_article() +
  scale_fill_manual(values = feature.colors, guide = guide_legend(reverse=TRUE)) +
  xlab("PC2") + ylab("SHAP values") + ggtitle("Sediment export reduction") +
  theme(axis.ticks.x = element_blank(), axis.text.x = element_blank(),
        legend.position = "none")

qbmarg_pc2_force <- ggplot(qbmarg_pc2, aes(rowid, value, fill = feature2)) +
  geom_col(width = 1) + egg::theme_article() +
  scale_fill_manual(values = feature.colors, guide = guide_legend(reverse=TRUE)) +
  xlab("PC2") + ylab("SHAP values") + ggtitle("Baseflow change") +
  theme(axis.ticks.x = element_blank(), axis.text.x = element_blank(),
        legend.position = "none")

marg_forcestackPC1 <- cowplot::plot_grid(polmarg_pc1_force, nexpmarg_pc1_force, align='v',
                                         labels = c("B","C"), ncol=1) %>% 
  ggplotGrob()
marg_forcestackPC2 <- cowplot::plot_grid(sedmarg_pc2_force, qbmarg_pc2_force, align='v',
                                         labels = c("D","E"), ncol=1) %>%
  ggplotGrob()

# make a legend

marglegend_dummy <- ggplotGrob(ggplot(polmarg_pc1, aes(rowid, value, fill = feature2)) +
  geom_col(width = 1) + egg::theme_article() +
  scale_fill_manual(values = feature.colors, 
                    guide = guide_legend(reverse=TRUE)) +
  theme(legend.position = "bottom", legend.title = element_blank()))$grobs

marg_legend <- marglegend_dummy[[which(sapply(marglegend_dummy, function(x)
      x$name) == "guide-box")]]

marg_plots <- cowplot::plot_grid(marg_forcestackPC1, marg_forcestackPC2, ncol = 2)

marg_plots_legend <- grid.arrange(marg_plots, marg_legend, 
                                  layout_matrix = 
                                    rbind(
                                      c(1,1,1,1,1,1),
                                      c(1,1,1,1,1,1),
                                      c(1,1,1,1,1,1),
                                      c(1,1,1,1,1,1),
                                      c(1,1,1,1,1,1),
                                      c(1,1,1,1,1,1),
                                      c(1,1,1,1,1,1),
                                      c(1,1,1,1,1,1),
                                      c(2,2,2,2,2,2)
                                    ))

marg_plots_all <- cowplot::plot_grid(biplot_marg_pca, marg_plots_legend, 
                                     labels = c("A",""), rel_widths = c(1,2.2))

# ggsave("Figs/marg_pcashap.png", marg_plots_all, width = 13, height = 5.5, bg = "white", dpi = 300)

```

```{r plot_pca_margoutcomes, fig.width = 12, fig.height = 5}
marg_plots_all
```