### 2025-04-04 Paper_analysis_percent.Rmd
another update to site

---
title: "Percent change"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r load}
library(tidyverse)
library(iml)
library(ranger)
load("data/ES-RF3pc.RDA")

```

## Ecosystem service improvements depend on the baseline condition

```{r data_setup}

dat.long <- dat %>% 
  pivot_longer(cols = c(1:8,13:16), names_to = "ES_response", values_to = "ES_value") %>%
  pivot_longer(cols = c("ntr_pland", "ann_agr_pland"), names_to = "Landuse",
               values_to = "Landuse_value") %>%
  select(ES_response, ES_value, Landuse, Landuse_value) %>%
  mutate(ES = gsub("delta_|_base|_(incr|red).margin|pc_","", ES_response)) %>%
  mutate(ES = case_match(ES,
                         "sed" ~ "sed_export",
                         "pol" ~ "pollinators",
                         "nexp" ~ "n_export",
                         .default = ES)) %>%
  mutate(response_type =
           gsub("sed|export|n_|pollinators|qb|_|pol|nexp|red.|incr.","",ES_response)) %>%
  mutate(response_type = case_match(
    response_type,
    "base" ~ "Baseline ES [/ha]",
    "delta" ~ "\u0394 ES [/ha]",
    "pcmargin" ~ "% \u0394 ES per ha margin"
  )) %>%
  mutate(response_type = factor(response_type, 
                                levels = c("Baseline ES [/ha]",
                                           "\u0394 ES [/ha]",
                                           "% \u0394 ES per ha margin")))

```

Ecosystem service improvements depend on the baseline condition, except for pollination.

This pattern holds even if we factor out margin area.

The trend is different if we look at % change in ecosystem services from baseline, per margin area

```{r percentchangedemo}

dat_wider <- dat.long %>% select(-ES_response) %>% 
  pivot_wider(names_from = response_type, values_from = ES_value)

base.v.margin <- ggplot(dat_wider, aes(`Baseline ES [/ha]`, `Δ ES [/ha]`)) + geom_point() + 
  xlab("Baseline ES per ha") + ylab("Increase in ES per ha margin (ES units)") +
  geom_smooth(method = "gam") + facet_wrap(~ES, scales = "free") + 
  ggtitle("ES baseline vs \nES improvement per margin") + egg::theme_article()

base.v.margin.base <- ggplot(dat_wider, aes(`Baseline ES [/ha]`, `% Δ ES per ha margin`)) + 
  geom_point() + geom_smooth(method = "gam") + facet_wrap(~ES, scales = "free") +
  ggtitle("ES Baseline vs \n% ES improvement per margin") + 
  xlab("Baseline ES per ha") +
  ylab("% improvement from baseline per ha margin") +
  egg::theme_article()

cowplot::plot_grid(base.v.margin, base.v.margin.base)

```

## Patterns with annual agriculture

Looking at ES % change from baseline shows some stronger patterns than before for sediment and water recharge.

```{r compplots, fig.width=9, fig.height=6}

# fit some gams
dat.gam <- dat.long %>% filter(Landuse == "ntr_pland") %>% 
  group_by(response_type, ES) %>%
  nest() %>%
  # Fit GAM
  mutate(Mod = map(data, ~mgcv::gam(ES_value ~ s(Landuse_value), data = .x))) %>%
  # Get R2
  mutate(R2 = map_dbl(Mod, ~round(summary(.x)$r.sq, 2)))

ggplot(dat.long %>% filter(Landuse == "ann_agr_pland"),
       aes(Landuse_value, ES_value)) + 
  geom_point() + geom_smooth(method="gam") +
  ggh4x::facet_grid2(response_type ~ ES,  scales = "free", independent = "y", 
                     switch = "y") + 
  egg::theme_article() + 
  # Add label
  geom_label(data = dat.gam, 
             aes(x = Inf, y = Inf, 
                 label = paste("R2 = ", R2, sep = " ")),
             hjust = 1, vjust = 1) +
  xlab("% Annual agriculture") +
  ylab("ES value") + theme(strip.placement = "outside")
```

## Random forest models for % improvement

Units for % improvement per margin ha are in tenths of a percent, so MSE looks smaller.

### Model performance
*Pollinators:*

     r2 = `r round(pollinators_pc.rf$r.squared, 3)`
     OOB MSE = `r format(round(pollinators_pc.rf$prediction.error, 5),scientific=F)`

*Nitrogen export:*

     r2 = `r round(n_export_pc.rf$r.squared, 3)`
     OOB MSE = `r format(round(n_export_pc.rf$prediction.error, 5),scientific=F)`

*Sediment export:*

     r2 = `r round(sed_pc.rf$r.squared, 3)`
     OOB MSE = `r format(round(sed_pc.rf$prediction.error, 5),scientific=F)`

*Baseflow:*

     r2 = `r round(qb_pc.rf$r.squared, 3)`
     OOB MSE = `r format(round(qb_pc.rf$prediction.error, 5),scientific=F)`

### Variable importance

```{r importance, fig.height=16, fig.width=9}
library(ggpattern)

metric.descr <- read.csv("data/lsm_description.csv")

imp.df <- bind_rows(
  pollinators_pc.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Pollinators", model = "base"),
  n_export_pc.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "N export", model = "base"),
  sed_pc.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Sed export", model = "base"),
  qb_pc.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Baseflow", model = "base")
) %>%
  mutate(sig = ifelse(pvalue <0.05, TRUE, FALSE)) %>%
  mutate(metric2 = tidytext::reorder_within(metric, importance,
                                           within = list(ES, model))) %>%
  left_join(metric.descr %>% select(lsm, type), by = c("metric"="lsm")) %>%
  mutate(type = case_match(
    metric,
    "field_pland" ~ "area and edge metric",
    "field_frac_mn" ~ "shape metric",
    "field_frac_cv" ~ "shape metric",
    "prn_pagr" ~ "area and edge metric",
    .default = type
  ))

# ggplot(imp.df, aes(metric2, importance, fill = type, pattern = sig)) + 
#   geom_bar_pattern(stat="identity",
#                    color = "black", 
#                    pattern_fill = "black",
#                    pattern_angle = 45,
#                    pattern_density = 0.1,
#                    pattern_spacing = 0.025,
#                    pattern_key_scale_factor = 0.6) + 
#   tidytext::scale_x_reordered() +
#   coord_flip() +
#   ggh4x::facet_grid2(ES~model, scales = "free", independent = "all") +
#   scale_pattern_manual(values = c("TRUE" = "none", "FALSE" = "stripe")) +
#   egg::theme_article() +
#   theme(legend.position = "top") + 
#   guides(pattern = guide_legend(override.aes = list(fill = "white")),
#          fill = guide_legend(override.aes = list(pattern = "none")))

```

```{r importance_shap}
load("data/ES-RF3pc-preds.RDA")   # prediction objects from iml package
load("data/ES-RF3pc-shapimportance.RDA")
library(shapviz)
library(gridExtra)

```

### Pollinators

#### Permutation based importance

```{r pol_permimport, fig.width=9}

ggplot(imp.df %>% filter(ES=="Pollinators"), aes(metric2, importance, fill = type, pattern = sig)) + 
  geom_bar_pattern(stat="identity",
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,
                   pattern_spacing = 0.025,
                   pattern_key_scale_factor = 0.6) + 
  tidytext::scale_x_reordered() +
  coord_flip() +
  scale_pattern_manual(values = c("TRUE" = "none", "FALSE" = "stripe")) +
  egg::theme_article() +
  theme(legend.position = "top") + 
  guides(pattern = guide_legend(override.aes = list(fill = "white")),
         fill = guide_legend(override.aes = list(pattern = "none")))

```

#### Shapley importance

```{r shap_pol, fig.width=9}
  sv_importance(ex.polpc.glb) + egg::theme_article()

```

#### Base pollinator model Shapley dependence plots

The two major contributors. Colors are the strongest interactor with that variable.

```{r shap_pol_baseshap, fig.width=9}
grid.arrange(
  sv_dependence(ex.polpc.glb, v="ntr_pland") + egg::theme_article(),
  sv_dependence(ex.polpc.glb, v="ann_agr_pland") + egg::theme_article(), 
  ncol = 2
)

```

**_Compare to PDP and ALE_**

```{r pol_base_ale, fig.width=9}

polbase.ntr_pland.pdp = FeatureEffect$new(pred.pollinators_pc, 
                                          feature = "ntr_pland",
                                          method = "pdp")
polbase.ann_agr_pland.pdp = FeatureEffect$new(pred.pollinators_pc, 
                                          feature = "ann_agr_pland",
                                          method = "pdp")

polbase.ntr_pland.ale = FeatureEffect$new(pred.pollinators_pc, 
                                          feature = "ntr_pland",
                                          method = "ale")
polbase.ann_agr_pland.ale = FeatureEffect$new(pred.pollinators_pc, 
                                          feature = "ann_agr_pland",
                                          method = "ale")

cowplot::plot_grid(
  polbase.ntr_pland.pdp$plot() + ggtitle("ntr_pland PDP") + egg::theme_article(),
  polbase.ntr_pland.ale$plot() + ggtitle("ntr_pland ALE") + egg::theme_article(),

  polbase.ann_agr_pland.pdp$plot() + ggtitle("ann_agr_pland PDP") + egg::theme_article(),
  polbase.ann_agr_pland.pdp$plot() + ggtitle("ann_agr_pland ALE") + egg::theme_article(),
  ncol = 2, align = "v"
)

```

### N Export

### Permutation based importance

```{r nexp_permimport, fig.width=9}

ggplot(imp.df %>% filter(ES=="N export"), aes(metric2, importance, fill = type, pattern = sig)) + 
  geom_bar_pattern(stat="identity",
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,
                   pattern_spacing = 0.025,
                   pattern_key_scale_factor = 0.6) + 
  tidytext::scale_x_reordered() +
  coord_flip() +
  scale_pattern_manual(values = c("TRUE" = "none", "FALSE" = "stripe")) +
  egg::theme_article() +
  theme(legend.position = "top") + 
  guides(pattern = guide_legend(override.aes = list(fill = "white")),
         fill = guide_legend(override.aes = list(pattern = "none")))

```

#### Shapley importance

```{r shap_nexp2, fig.width=9}

  sv_importance(ex.nexppc.glb) + egg::theme_article()

```

#### Base N export shapley dependence plots

Two major -ish contributors.

```{r shap_nexp_pcshap, fig.width=9}

cowplot::plot_grid(
  sv_dependence(ex.nexppc.glb, v="ntr_pland") + egg::theme_article(),
  sv_dependence(ex.nexppc.glb, v="ntr_cohesion") + egg::theme_article(), 
  ncol = 2, align = "v"
)

```

**_Compare to PDP and ALE_**

```{r nexp_pc_ale, fig.width=9}

nexppc.ntr_pland.pdp = FeatureEffect$new(pred.nexp_pc, 
                                          feature = "ntr_pland",
                                          method = "pdp")
nexppc.ntr_cohesion.pdp = FeatureEffect$new(pred.nexp_pc, 
                                          feature = "ntr_cohesion",
                                          method = "pdp")

nexppc.ntr_pland.ale = FeatureEffect$new(pred.nexp_pc, 
                                          feature = "ntr_pland",
                                          method = "ale")
nexppc.ntr_cohesion.ale = FeatureEffect$new(pred.nexp_pc, 
                                          feature = "ntr_cohesion",
                                          method = "ale")

cowplot::plot_grid(
  nexppc.ntr_pland.pdp$plot() + ggtitle("ntr_pland PDP") + egg::theme_article(),
  nexppc.ntr_pland.ale$plot() + ggtitle("ntr_pland ALE") + egg::theme_article(),

  nexppc.ntr_cohesion.pdp$plot() + ggtitle("ntr_cohesion PDP") + egg::theme_article(),
  nexppc.ntr_cohesion.ale$plot() + ggtitle("ntr_cohesion ALE") + egg::theme_article(),

  ncol = 2, align = "v"
)

```

### Sediment Export

### Permutation based importance

```{r sed_permimport, fig.width=9}

ggplot(imp.df %>% filter(ES=="Sed export"), aes(metric2, importance, fill = type, pattern = sig)) + 
  geom_bar_pattern(stat="identity",
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,
                   pattern_spacing = 0.025,
                   pattern_key_scale_factor = 0.6) + 
  tidytext::scale_x_reordered() +
  coord_flip() +
  scale_pattern_manual(values = c("TRUE" = "none", "FALSE" = "stripe")) +
  egg::theme_article() +
  theme(legend.position = "top") + 
  guides(pattern = guide_legend(override.aes = list(fill = "white")),
         fill = guide_legend(override.aes = list(pattern = "none")))

```

#### Shapley importance

```{r shap_sed, fig.width=9}

sv_importance(ex.sedpc.glb) + egg::theme_article()

```

#### Percent improvement in sediment export Shapley dependence plots

One major contributor

```{r shap_sed_baseshap, fig.width=9}
sv_dependence(ex.sedpc.glb, v="ntr_pland") + egg::theme_article()

```

**_Compare to PDP and ALE_**

```{r sedexp_pc_ale, fig.width=9}

sedpc.ntr_pland.pdp = FeatureEffect$new(pred.sed_pc, 
                                          feature = "ntr_pland",
                                          method = "pdp")

sedpc.ntr_pland.ale = FeatureEffect$new(pred.sed_pc, 
                                          feature = "ntr_pland",
                                          method = "ale")

cowplot::plot_grid(
  sedpc.ntr_pland.pdp$plot() + ggtitle("ann_agr_ed PDP") +
    egg::theme_article(),
  sedpc.ntr_pland.ale$plot() + ggtitle("ann_agr_ed ALE") +
    egg::theme_article(),

  ncol = 2, align = "v"
)

```

### Water recharge

### Permutation based importance

```{r qb_permimport, fig.width=9}

ggplot(imp.df %>% filter(ES=="Baseflow"), aes(metric2, importance, fill = type, pattern = sig)) + 
  geom_bar_pattern(stat="identity",
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,
                   pattern_spacing = 0.025,
                   pattern_key_scale_factor = 0.6) + 
  tidytext::scale_x_reordered() +
  coord_flip() +
  scale_pattern_manual(values = c("TRUE" = "none", "FALSE" = "stripe")) +
  egg::theme_article() +
  theme(legend.position = "top") + 
  guides(pattern = guide_legend(override.aes = list(fill = "white")),
         fill = guide_legend(override.aes = list(pattern = "none")))

```

#### Shapley importance

```{r shap_qb, fig.width=9}

sv_importance(ex.qbpc.glb) + egg::theme_article()

```

#### Base water runoff Shapley depednence plot

One major contributor.

```{r shap_qb_baseshap, fig.width=9}
cowplot::plot_grid(
  sv_dependence(ex.qbpc.glb, v="ntr_ed") + egg::theme_article(),
  sv_dependence(ex.qbpc.glb, v="prn_agr_pd") + egg::theme_article(),
  ncol = 2, align = "v"
)

```

**_Compare to PDP and ALE_**

```{r qb_pc_ale, fig.width=9}

qbpc.ntr_ed.pdp = FeatureEffect$new(pred.qb_pc, 
                                             feature = "ntr_ed",
                                             method = "pdp")

qbpc.ntr_ed.ale = FeatureEffect$new(pred.qb_pc, 
                                              feature = "ntr_ed",
                                              method = "ale")

qbpc.prn_agr_pd.pdp = FeatureEffect$new(pred.qb_pc, 
                                             feature = "prn_agr_pd",
                                             method = "pdp")

qbpc.prn_agr_pd.ale = FeatureEffect$new(pred.qb_pc, 
                                              feature = "prn_agr_pd",
                                              method = "ale")

cowplot::plot_grid(
  qbpc.ntr_ed.pdp$plot() + ggtitle("ntr_ed PDP") +
    egg::theme_article(),
  qbpc.ntr_ed.ale$plot() + ggtitle("ntr_ed ALE") +
    egg::theme_article(),

  qbpc.prn_agr_pd.pdp$plot() + ggtitle("prn_agr_pd PDP") +
    egg::theme_article(),
  qbpc.prn_agr_pd.ale$plot() + ggtitle("prn_agr_pd ALE") +
    egg::theme_article(),

  ncol = 2, align = "v"
)

```